---
title: "Soil AWS descriptive"
author: "Maite"
date: "3 12 2021"
output: html_document
---

```{r setup, include=FALSE, message=F, warning=F}
knitr::opts_chunk$set(echo = FALSE, message=F, warning=F)
```

```{r}
library(here)
library(readxl)
library(kableExtra)
library(dplyr)
library(ggplot2)
library(padr)
library(reshape2)
```

# Dataset for all stations

```{r}
saws <- readRDS(here("../calib_valid/AWS/SoilStations/all_AWSS_all_years.rds"))

var_together <- data.frame(read_xlsx(here("../calib_valid/AWS/SoilStations/new_vs_old_colnames.xlsx"),
                                     sheet=2))

kbl(var_together[,c(1:3)],
    caption="All variables the stations have in common") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = F, 
                position = "left")
```


```{r}
# summary(saws)
kbl(head(saws)) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = F, 
                position = "left")


tail(sort(saws$SMC))
ol <- which(saws$SMC > 30,) # not sure if that is justified
# getting rid of those outliers
saws <- saws[-ol, ]

hist(log((saws$SMC)+1), breaks=90)
tail(sort(saws$SMC))

```


# Dataset to check for relation of RH and SMC
```{r}
all_useful_var <- data.frame(read_xlsx(here("../calib_valid/AWS/SoilStations/new_vs_old_colnames.xlsx"),
                                       sheet=1))

kbl(all_useful_var,
    caption="All useful variables") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = F, 
                position = "left",
                font_size = 9)
```

Only possible to look into BP and BPE, DJP also has a soil moisture water 
measurement at 2cm height, but it is only 0 and in a few cases we've got crazy outliers.. 
doesn't really make sense.
```{r}
stations <- readRDS(here("../calib_valid/AWS/SoilStations/stations_before_unifying.rds"))

srh <- list(BP=stations$BP,
     BPE=stations$BPE)

RH_SMC_T_Plots <- lapply(seq(srh), function(s){
      if(s==1){
              rhT <- data.frame(RH=srh[[s]]$`Soil_RH_2_cm_%`,
                          SMC=srh[[s]]$`WATER_2_cm,_1_wfv`,
                          Temp=srh[[s]]$Soil_Temp_2_cm_deg_C, 
                          date=srh[[s]]$DATE_NA_NA)
      } else if (s==2){
              rhT <- data.frame(RH=srh[[s]]$`Soil_RH_2_cm_%`,
                  SMC=srh[[s]]$`WATER_2_cm,_1_wfv`,
                  Temp=srh[[s]]$Soil_Temp_2_cm_deg_C, 
                  date=srh[[s]]$DATE_NA_NA)
      } 
      rhT$Temp[rhT$Temp=="NAN"] <- NA
      rhT$Temp <- as.numeric(rhT$Temp)
      rhT$Temp[rhT$Temp< (-100)] <- NA # one outlier at -200°C
      rhT$RH[rhT$RH == -100] <- NA # assuming this means no data 
      
      rhT <- rhT[complete.cases(rhT),]
      rhT$log_SMC_1 <- log(rhT$SMC+1)

      rhT <- pad(rhT) # fill up with NA to have gaps in linegraph 
      
      scaleFactor <-  max(rhT$RH, na.rm=T) / max(rhT$log_SMC_1, na.rm=T)
      scaleFactorT <- (max(rhT$RH, na.rm=T) - min(rhT$RH, na.rm=T)) / 
                      (max(rhT$Temp, na.rm=T) - min(rhT$Temp, na.rm=T))
      
      add <- mean(rhT$RH,na.rm=T)-mean(rhT$Temp,na.rm=T)

      RHcolor <-"#50C7C7"
      Tcolor <- "#C75067"
      SMCcolor <- "#5071C7"
      
      RH_SMC_P <- ggplot(rhT, aes(x=date)) +
      
            geom_line(aes(y=log_SMC_1), color=SMCcolor)+
            geom_point(aes(y=log_SMC_1), color=SMCcolor, size=1)+
            geom_line(aes(y=RH/ scaleFactor), color=RHcolor)+
            geom_point(aes(y=RH/scaleFactor), color=RHcolor, size=1)+
      
            scale_y_continuous(
              
              # Features of the first axis
              name = "  log SMC (wfv)",
              
              # Add a second axis and specify its features
              sec.axis = sec_axis(~.*scaleFactor, name="RH (%)")
              
              
            ) +theme_minimal()+
            theme(axis.title.y.left=element_text(color=SMCcolor,size=12, face="bold"),
                axis.text.y.left=element_text(color=SMCcolor,size=12),
                axis.title.y.right=element_text(color=RHcolor,size=12, face="bold"),
                axis.text.y.right=element_text(color=RHcolor, size=12))+
            ggtitle("Relative Humidity and Soil Moisture Content (wfv)")
      
      T_P <- ggplot(rhT, aes(x=date))+
            geom_line(aes(y=Temp), color=Tcolor)+
            geom_point(aes(y=Temp), color=Tcolor, size=1)+
            geom_line(aes(y=RH/ scaleFactorT-add), color=RHcolor)+
            geom_point(aes(y=RH/scaleFactorT-add), color=RHcolor, size=1)+
      
            scale_y_continuous(
              
              # Features of the first axis
              name = " Temperature (°C)",
              
              # Add a second axis and specify its features
              sec.axis = sec_axis(~.*scaleFactorT+add, name="RH (%)")

            )+
            theme_minimal()+
            theme(axis.title.y.left=element_text(color=Tcolor,size=12, face="bold"),
                  axis.text.y.left=element_text(color=Tcolor,size=12),
                  axis.title.y.right=element_text(color=RHcolor,size=12, face="bold"),
                  axis.text.y.right=element_text(color=RHcolor, size=12))+
            ggtitle("Temperature")
      
      p <- gridExtra::arrangeGrob(RH_SMC_P, T_P, nrow = 2)
      return(p)
          
})

```

Log transformation to see better what is happening with SMC:
log(SMC+1) because that way 0 doesn't result in -Inf. 
Where SMC is 0, log(SMC+1) is 0 too. 
Highest measured SMC is 0.165, which  translates to 0.152 log(SMC+1). 


## Bull Pass 
```{r}
grid::grid.draw(RH_SMC_T_Plots[[1]])
```

## Bull Pass East 
```{r}
grid::grid.draw(RH_SMC_T_Plots[[2]])
```




```{r, warning=FALSE, message=FALSE, eval=F}
  
# make an overview of names to find out which are out of line 
nams <- lapply(seq(length(f)), function(y){
    names(stat_year[[y]])
})
namdf <- do.call(rbind.data.frame, nams)
names(namdf) <- paste0("V", seq(names(namdf)))
# View(namdf)
names(stat_year) <- sapply(seq(stat_year), function(i){ as.character(stat_year[[i]][1,2]) })

# sort out 2009
stat_year_clean <- stat_year[names(stat_year)!="2009"]
# length(stat_year_clean)
nams <- lapply(seq(length(stat_year_clean)), function(y){
    names(stat_year_clean[[y]])
})
namdf <- do.call(rbind.data.frame, nams)
names(namdf) <- paste0("V", seq(names(namdf)))
# View(namdf)
# any(namdf[1,] != namdf[4,])

bullpass <- do.call("rbind.data.frame", stat_year_clean)

names(bullpass) <- gsub(",", "_", names(bullpass))

names(bullpass)[104]
names(bullpass)[which(names(bullpass)=="SOIL_RH_%")[1]] <- "SOIL_RH_%_1"
names(bullpass)[which(names(bullpass)=="SOIL_ELECT_RES")[1]] <- "SOIL_ELECT_RES_1"

# filling the gaps in incomplete datetime variables so that no 
#data are actually no data in the plot
bullpass <- pad(bullpass) 

bullpass$useable <- NA
bullpass$useable[bullpass$Soil_Temp_2_cm_White_deg_C < (-1)] <- 0
bullpass$useable <- as.factor(bullpass$useable)
```


```{r, warning=FALSE, message=FALSE, eval=F}

prh <- ggplot(bullpass, aes(DATE_NA_NA, WATER_2_cm__1_wfv))+
  geom_point(size=0.5)+
  geom_line()+
  ggtitle("Relative Humidity")+
  theme_minimal()
pst <- ggplot(bullpass, aes(DATE_NA_NA, Soil_Temp_2_cm_White_deg_C))+
  geom_point(size=0.5)+
  geom_line()+
  geom_hline(yintercept = -1, lty="dashed")+
  ggtitle("Soil temperature at surface level")+
  theme_minimal()

gridExtra::grid.arrange(prh,pst)


prh <- ggplot(bullpass, aes(DATE_NA_NA, WATER_2_cm__1_wfv, color=useable))+
  geom_point(size=0.5)+
  scale_color_manual(values=c("#e1e2e6","#0d0d0c"))+
  ggtitle("Relative Humidity")+
  theme_minimal()
pst <- ggplot(bullpass, aes(DATE_NA_NA, Soil_Temp_2_cm_White_deg_C, color=useable))+
  geom_point(size=0.5)+
  geom_hline(yintercept = -1, lty="dashed")+
  scale_color_manual(values=c("#e1e2e6","#0d0d0c"))+
  ggtitle("Soil temperature at surface level")+
  theme_minimal()

gridExtra::grid.arrange(prh,pst)
```

