---
title: "Calibration and validation data for surface moisture in the McMurdo Dry Valleys"
author: "Maite"
date: "13 9 2021"

output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning=F, echo = F)
```


```{r}
library(sf)
library(ggplot2)
library(gridExtra)
library(grid)
library(colorspace)
library(here)
library(dplyr)
library(lubridate)
library(padr)
```

This skript shows calibration and validation data for surface moisture modelling in the McMurdo Dry Valleys in a descriptive way. 

# iButton data

Read in iButton data for the different valleys. 

```{r, message=F, results='hide'}
iButDir <- paste0(here("data/calib_valid/iButtons/"))

gpkdir <- paste0(iButDir, "/gpkg/")
f <- list.files(gpkdir,full.names=T)

iB_valley <- lapply(seq(f),function(i){
  st_read(f[[i]])
})


```

Project iButton data to EPSG 3031
```{r}
predDir <- here("data/predictors/") 
predcrs <- readRDS(list.files(predDir, pattern="predcrs", full.names=T)) 

iB_valley <- lapply(seq(iB_valley), function(i){
  x <- st_transform(iB_valley[[i]], crs = st_crs(predcrs$input))
  x$Site_Name <- factor(x$Site_Name)
  x
})
names(iB_valley) <- tools::file_path_sans_ext(basename(f))
```


Use only data, where temperature is higher than -1°C
```{r}

onlySummer <- TRUE 

unique_locations <- st_difference(iB_valley$iBut)
valleys <- unique(unique_locations$Valley) 

iButPlots <- lapply(seq(valleys), function(v){
  iBut_Vlly <- iB_valley[[which(names(iB_valley)==valleys[v])]]
  iBut_Vlly <- iBut_Vlly[order(iBut_Vlly$time),]
  # cut into two to not get the gap in the middle where temperatures are too low
  timeslot1 <- iBut_Vlly[order(iBut_Vlly$time) < (max(order(iBut_Vlly$time))/2),] 
  timeslot2 <- iBut_Vlly[order(iBut_Vlly$time) > (max(order(iBut_Vlly$time))/2),] 
  
  iBut_Vlly <- list(timeslot1, timeslot2)
  
  fp <- lapply(seq(iBut_Vlly), function(i){
    
      if(onlySummer == TRUE){
        iBut_Vlly[[i]] <- iBut_Vlly[[i]][month(iBut_Vlly[[i]]$time) %in% c(11,12,1,2),] 
      }
      
      iBut_Vlly[[i]]$Rh[iBut_Vlly[[i]]$T < -1] <- NA
      
      fs <- unique(year(iBut_Vlly[[i]]$time))
      
      pRh <- ggplot(iBut_Vlly[[i]], aes(x=time, y=Rh, color=Site_Name)) +
            scale_colour_viridis_d()+
            geom_line() +
            ggtitle(paste0("field season ", paste(fs, collapse = '/')))+
            xlab("")
    
    
      pT <- ggplot(iBut_Vlly[[i]], aes(x=time, y=T, color=Site_Name)) +
            scale_colour_viridis_d()+
            geom_line() +
            geom_hline(yintercept = -1, linetype = "dotted")+
            xlab("")
      
      fp <- arrangeGrob(pRh, pT, nrow = 2)
  })
  return(fp)  

})

```


To see the iButton logger location names, hold cursor over point.

### Hidden_Valley
```{r}
grid.draw(iButPlots[[1]][[1]])
grid.draw(iButPlots[[1]][[2]])

```


### Wright_Valley  
```{r}
grid.draw(iButPlots[[2]][[1]])
grid.draw(iButPlots[[2]][[2]])

```

### McKelvey_Valley
```{r}
grid.draw(iButPlots[[3]][[1]])
grid.draw(iButPlots[[3]][[2]])

```

### Victoria_Valley
```{r}
grid.draw(iButPlots[[4]][[1]])
grid.draw(iButPlots[[4]][[2]])

```

### Alatna_Valley  
```{r}
grid.draw(iButPlots[[5]][[1]])
grid.draw(iButPlots[[5]][[2]])

```

### Taylor_Valley
```{r}
grid.draw(iButPlots[[6]][[1]])
grid.draw(iButPlots[[6]][[2]])

```

TV-S-500 seems to be below a snow patch in a shady corner and is fully saturated. 

# Automatic Weather Station Data
### Lake Fryxell Met Station
Dataset can be downloaded here: [Lake Fryxell](https://mcm.lternet.edu/content/high-frequency-measurements-lake-fryxell-meteorological-station-frlm-mcmurdo-dry-valleys)

There are two datasets, one for relative humidity and one for soil moisture,
they are structured as follows: 

```{r}
awsDir <- here("data/calib_valid/AWS/")
lf_rh <- list.files(awsDir, full.names=T, pattern="frlm_rh")
lf_st <- list.files(awsDir, full.names=T, pattern="frlm_soilt")

f_rh <- read.csv(lf_rh)
head(f_rh, n=3)
f_rh$datetime <- as.POSIXct(f_rh$DATE_TIME, format="%d/%m/%Y %H:%M")

f_st <- read.csv(lf_st)
head(f_st, n=3)
f_st$datetime <- as.POSIXct(f_st$DATE_TIME, format="%d/%m/%Y %H:%M")
```

Years that are covered in this dataset are: 
```{r, out.width = '40%'}
hist(year(f_rh$datetime), main="RH Observations per year", 
     xlab="Year", ylab="n observations")
hist(year(f_st$datetime), main="Soil Temperature Observations per year", 
     xlab="Year", ylab="n observations")
```

Let's for now take a look at the year 2013 and focus on Relative Humidity and 
the soil temperature at surface level. We're excluding timestamps where surface 
soil temperature is below -1°C. When soil and surface water freezes no reliable 
relative humidity information can be measured any more. 
```{r}
f_rh13 <- f_rh[year(f_rh$datetime)==2013,]
f_st13 <- f_st[year(f_st$datetime)==2013,]

# filling the gaps in incomplete datetime variables so that no 
#data are actually no data in the plot
f_rh13 <- pad(f_rh13) 
f_st13 <- pad(f_st13)

# which timestamps are useable? 
f_rh13$useable <- 1
f_rh13$useable[f_st13$SOILT0CM < (-1)] <- 0
f_rh13$useable <- factor(f_rh13$useable)

f_st13$useable <- 1
f_st13$useable[f_st13$SOILT0CM < (-1)] <- 0
f_st13$useable <- factor(f_st13$useable)

prh <- ggplot(f_rh13, aes(datetime, RH, color=useable))+
  geom_point(size=0.5)+
  scale_color_manual(values=c("#e1e2e6","#0d0d0c"))+
  ggtitle("Relative Humidity")+
  theme_minimal()
pst <- ggplot(f_st13, aes(datetime, SOILT0CM, color=useable))+
  geom_point(size=0.5)+
  geom_hline(yintercept = -1, lty="dashed")+
  scale_color_manual(values=c("#e1e2e6","#0d0d0c"))+
  ggtitle("Soil temperature at surface level")+
  theme_minimal()

gridExtra::grid.arrange(prh,pst)


```

