---
title: "Calibration and validation data for surface moisture in the McMurdo Dry Valleys"
author: "Maite"
date: "13 9 2021"

output: 
 rmarkdown::html_document:
  #bookdown::html_document2:
    theme: flatly
    highlight: tango
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 6
link-citations: yes
toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning=F, echo = F)
```


```{r}
library(sf)
library(mapview)
library(ggplot2)
library(gridExtra)
library(grid)
library(colorspace)
library(here)
library(dplyr)
library(lubridate)
```

This skript shows calibration and validation data for surface moisture modelling in the McMurdo Dry Valleys in a descriptive way. 

# iButton data

Read in iButton data for the different valleys. 

```{r, message=F, results='hide'}
iButDir <- paste0(here("data/calib_valid/iButtons/"))

gpkdir <- paste0(iButDir, "/gpkg/")
f <- list.files(gpkdir,full.names=T)

iB_valley <- lapply(seq(f),function(i){
  st_read(f[[i]])
})


```

<!-- Get DEM to write elevation into the location data  -->
<!-- ```{r} -->
<!-- predDir <- here("data/predictors/") -->
<!-- pred <- readRDS(list.files(predDir, pattern="full", full.names=T)) -->

<!-- pred <- pred[[c("dem", "slope", "aspect", -->
<!--        "soilraster", "landcoverres")]] -->
<!-- predcrs <- readRDS(list.files(predDir, pattern="predcrs", full.names=T)) -->
<!-- dem <- raster::raster(list.files(predDir,                                pattern="DEM_8m_MDV_clean_aoi_filled_filtered_new.tif", full.names=T)) -->
<!-- ``` -->

Project iButton data to EPSG 3031
```{r}
predDir <- here("data/predictors/") 
predcrs <- readRDS(list.files(predDir, pattern="predcrs", full.names=T)) 
iB_valley <- lapply(seq(iB_valley), function(i){
  x <- st_transform(iB_valley[[i]], crs = predcrs)
  x$Site_Name <- factor(x$Site_Name)
  x
})
names(iB_valley) <- tools::file_path_sans_ext(basename(f))
```


<!-- ```{r} -->
<!-- # How many locations per valley?  -->
<!-- loc_p_v <- sapply(seq(iB_valley), function(i){ -->
<!--   length(unique(iB_valley[[i]]$geom)) -->
<!-- }) -->
<!-- # max = 22 for each unique valley -->
<!-- data.frame("valley"=names(iB_valley), "n_locations"=loc_p_v) -->

<!-- # get locations of unqiue Site_Names -->
<!-- unique_locations <- st_difference(iB_valley$iBut) -->

<!-- pred_loc <- raster::extract(pred, unique_locations) -->

<!-- unique_locations <- cbind(unique_locations, pred_loc) -->

<!-- ``` -->



<!-- ## Display locations  -->
<!-- ```{r, message=F} -->
<!-- # i=1 -->
<!-- # cat("unique sensor locations: ", length(unique(iB_valley[[i]]$geom))) -->
<!-- # l <- length(unique(iB_valley[[i]]$geom)) -->
<!-- # unique(iB_valley[[i]]$Site_Name) -->

<!-- valleys <- unique(unique_locations$Valley) -->

<!-- plots <- lapply(seq(valleys), function(v){ -->
<!--   # cat(valleys[v], ":") -->
<!--   iBut_Vlly <- unique_locations[unique_locations$Valley==valleys[v],] -->
<!--   ex <- st_buffer(st_as_sfc(st_bbox(iBut_Vlly)),dist=200) -->
<!--   DEM_Vlly <- raster::crop(dem,y=st_bbox(ex)) -->
<!--   mapview(DEM_Vlly,na.color="#00000000", -->
<!--           map.types = "Esri.WorldImagery")+mapview(iBut_Vlly,  -->
<!--                                                    label="Site_Name",zcol="dem")  -->
<!-- }) -->
<!-- ``` -->

Use only data, where temperature is higher than -1°C
```{r}

onlySummer <- TRUE 

unique_locations <- st_difference(iB_valley$iBut)
valleys <- unique(unique_locations$Valley) 

iButPlots <- lapply(seq(valleys), function(v){
  iBut_Vlly <- iB_valley[[which(names(iB_valley)==valleys[v])]]
  iBut_Vlly <- iBut_Vlly[order(iBut_Vlly$time),]
  # cut into two to not get the gap in the middle where temperatures are too low
  timeslot1 <- iBut_Vlly[order(iBut_Vlly$time) < (max(order(iBut_Vlly$time))/2),] 
  timeslot2 <- iBut_Vlly[order(iBut_Vlly$time) > (max(order(iBut_Vlly$time))/2),] 
  
  iBut_Vlly <- list(timeslot1, timeslot2)
  
  fp <- lapply(seq(iBut_Vlly), function(i){
    
      if(onlySummer == TRUE){
        iBut_Vlly[[i]] <- iBut_Vlly[[i]][month(iBut_Vlly[[i]]$time) %in% c(11,12,1,2),] 
      }
      
      iBut_Vlly[[i]]$Rh[iBut_Vlly[[i]]$T < -1] <- NA
    
      pRh <- ggplot(iBut_Vlly[[i]], aes(x=time, y=Rh, color=Site_Name)) +
            scale_colour_viridis_d()+
            geom_line() +
            xlab("")
    
    
      pT <- ggplot(iBut_Vlly[[i]], aes(x=time, y=T, color=Site_Name)) +
            scale_colour_viridis_d()+
            geom_line() +
            geom_hline(yintercept = -1, linetype = "dotted")+
            xlab("")
      
      fp <- arrangeGrob(pRh, pT, nrow = 2)
  })
  return(fp)  

})

```


To see the iButton logger location names, hold cursor over point.

### Hidden_Valley
```{r}
# plots[[1]]
grid.draw(iButPlots[[1]][[1]])
grid.draw(iButPlots[[1]][[2]])

```


### Wright_Valley  
```{r}
# plots[[2]]
grid.draw(iButPlots[[2]][[1]])
grid.draw(iButPlots[[2]][[2]])

```

### McKelvey_Valley
```{r}
# plots[[3]]
grid.draw(iButPlots[[3]][[1]])
grid.draw(iButPlots[[3]][[2]])

```

### Victoria_Valley
```{r}
# plots[[4]]
grid.draw(iButPlots[[4]][[1]])
grid.draw(iButPlots[[4]][[2]])

```

### Alatna_Valley  
```{r}
# plots[[5]]
grid.draw(iButPlots[[5]][[1]])
grid.draw(iButPlots[[5]][[2]])

```

### Taylor_Valley
```{r}
# plots[[6]]
grid.draw(iButPlots[[6]][[1]])
grid.draw(iButPlots[[6]][[2]])

```

TV-S-500 seems to be below a snow patch in a shady corner and is fully saturated. 


# Automatic Weather Station Data
### Lake Fryxell Met Station
Dataset can be downloaded here: [Lake Fryxell](https://mcm.lternet.edu/content/high-frequency-measurements-lake-fryxell-meteorological-station-frlm-mcmurdo-dry-valleys)


```{r}
awsDir <- here("data/calib_valid/AWS/")
lf_rh <- list.files(awsDir, full.names=T, pattern="frlm_rh")
lf_st <- list.files(awsDir, full.names=T, pattern="frlm_soilt")

f_rh <- read.csv(lf_rh)
head(f_rh, n=3)
# str(f_rh)
# unique(f_rh$RH_COMMENTS)
f_rh$datetime <- as.POSIXct(f_rh$DATE_TIME, format="%d/%m/%Y %H:%M")

f_st <- read.csv(lf_st)
head(f_st, n=3)
# str(f_st)
# unique(f_st$SOILT_COMMENTS)
f_st$datetime <- as.POSIXct(f_st$DATE_TIME, format="%d/%m/%Y %H:%M")
# unique(year(f_st$datetime))
```

Years that are covered in this dataset are: 
```{r}
hist(year(f_rh$datetime), main="RH Observations per year", 
     xlab="Year", ylab="n observations")
hist(year(f_st$datetime), main="Soil Temperature Observations per year", 
     xlab="Year", ylab="n observations")
```


Let's for now take a look at the year 2019 and focus on Relative Humidity and the soil temperature at surface level. 
```{r}
f_rh19 <- f_rh[year(f_rh$datetime)==2019,]
f_st19 <- f_st[year(f_st$datetime)==2019,]

prh <- ggplot(f_rh19, aes(datetime, RH))+
  geom_line()+
  ggtitle("Relative Humidity")
pst <- ggplot(f_st19, aes(datetime, SOILT0CM))+
  geom_line()+
  ggtitle("Soil temperature at surface level")

gridExtra::grid.arrange(prh,pst)
```

