---
title: "get Landsat data"
author: "Maite"
date: "13 12 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message=F, warning=F)
```

## Landsat

* VIS
* SWIR 
    * Landsat 8 launched Feb 2013, 
    * Landsat 7 bands 5 and 7 similar to SWIR in L8, since 1999, sensor noise issue, 
    * Landsat 4-5 bands as in 7 since 1984-2013
        
```{r}
library(getSpatialData)
library(raster)
library(here)
library(rgdal)
library(ggplot2)
library(sf)
library(espa.tools)
```


```{r}
source(here("../login.R"))
# services()
```

## Basic settings 
```{r}
aoi <- readOGR(here("../aux_data/Levy_MDV_actually.shp"))
tr_preStudy <- c("2011-01-01", "2014-01-01")
#  archive directory:
Ldir <- "D:/Surface_Moisture/Landsat_data/"
set_archive(Ldir)

# area of interest (AOI):
# Use the example AOI or draw an AOI by calling set_aoi():
set_aoi(aoi)
# View the AOI:
view_aoi()

# get_products()
Ldir <- "D:/Surface_Moisture/Landsat_data/"
downloadpathL8 <- paste0(Ldir, "espa_order_downloads_L8_new/")
downloadpathL7 <- paste0(Ldir, "espa_order_downloads_L7_new/")

```

## Query for available datasets 

#### Only T2 available 
Landsat scenes not meeting Tier 1 criteria during processing are assigned to Tier 2. Tier 2 scenes adhere to the same radiometric standard as Tier 1 scenes, but do not meet the Tier 1 geometry specification due to less accurate orbital information (specific to older Landsat sensors), significant cloud cover, insufficient ground control, or other factors. 

```{r, eval=F}
# Landsat_records <- get_records(time_range = tr_preStudy,
#                        products = c("landsat_8_c1"))
# Landsat_records_LSR <- get_records(time_range = tr_preStudy,
#                                products="lsr_landsat_8_c1") # sr and bt 
# Landsat_records_OT_L2 <- get_records(time_range = tr_preStudy, # contains indices 
#                                products="landsat_ot_c2_l2")
# Landsat_records_ETM_L2 <- get_records(time_range = tr_preStudy,
#                                products="landsat_etm_c2_l2")
```

Collection 2 seems to contain only indices, in Col 1 there is surface reflectance (sr), brightness temperature (bt)  data available. 
The LSR seems to be reducing data only slightly (3220 vs. 3248 scenes)
I.e. use collection 1, lsr product 
```{r, eval=F}
# unique(Landsat_records_OT_L2$level)
# unique(Landsat_records_ETM_L2$level)
# unique(Landsat_records_LSR$level)
# unique(Landsat_records$level)
# 
# nrow(Landsat_records)
# nrow(Landsat_records_LSR)
# 
# unique(Landsat_records_LSR$collection_category)
# unique(Landsat_records$collection_category)

```

## Subset L8 records
```{r, eval=F}
L_records <- get_records(time_range = tr_preStudy,
                               products="lsr_landsat_8_c1") # sr and bt 

```

select brightness temperature, top of atmosphere and surface reflectance,
cloud cover < 30, overlap with aoi at least xxx,
check availability, order and save query for download via espa order. 
```{r, eval=F}
records <- L_records[L_records$level == "bt" | L_records$level == "sr",]

# calculate overlap from footprints

calc_footprint_aoi_overlap <- function(aoi, prop_aoi, query){
  fp <- query$footprint
  does_is <- st_intersects(fp, st_as_sf(aoi), sparse = F)
  query$intersects <- does_is
  is <- st_intersection(fp, st_as_sf(aoi))
  area_is <- st_area(is)
  aoiarea <- st_area(st_as_sf(aoi))
  aoiprop <- area_is/aoiarea
  query$aoiprop[query$intersects==TRUE] <- aoiprop
  sel <- as.numeric(aoiprop) >= prop_aoi
  query$sel_overlap <- NA
  query$sel_overlap[query$intersects==TRUE] <- sel
  return(query)
}

aoiwgs <- spTransform(aoi, CRSobj = CRS("+init=epsg:4326"))
records <- calc_footprint_aoi_overlap(aoi=aoiwgs, 
                                      prop_aoi = 0.3, 
                                      query=records)
# records$product_group <- "landsat"

records <- check_availability(records, verbose=TRUE) # this is not working, thus next line
# records$download_available <- FALSE
table(records$download_available)

# add summary for ordering 
records$summary <- paste0("ID: ", records$record_id, " Acquisition Date: ", records$date_acquisition,
" Path: ", records$tile_number_horizontal , " Row: ", records$tile_number_vertical)

records <- records[records$cloudcov_land < 30,] # cloud cover 
nrow(records)/2

records <- records[records$sel_overlap==TRUE,]

records <- order_data(records)


records$download_available
records$order_status


saveRDS(records, paste0(Ldir, "records_new.RDS"))

# edit(getAnywhere('order_data'), file='source_orer_data.r')


```


## Subset L7 records
```{r, eval=F}
L7_records <- get_records(time_range = tr_preStudy,
                               products="lsr_landsat_etm_c1") # sr and bt 

```

select brightness temperature, top of atmosphere and surface reflectance,
cloud cover < 30, check availability, order and download
```{r, eval=F}
L7_records <- L7_records[L7_records$level == "bt" | L7_records$level == "sr",]
L7_records <- check_availability(L7_records)
table(L7_records$download_available)
# add summary for ordering 
L7_records$summary <- paste0("ID: ", L7_records$record_id, " Acquisition Date: ", L7_records$date_acquisition,
" Path: ", L7_records$tile_number_horizontal , " Row: ", L7_records$tile_number_vertical)

L7_records <- L7_records[L7_records$cloudcov_land < 30,] # cloud cover 

L7_records <- calc_footprint_aoi_overlap(aoi=aoiwgs, 
                                      prop_aoi = 0.3, 
                                      query=L7_records)
L7_records <- L7_records[L7_records$sel_overlap==TRUE,]

# records$product_group <- "landsat"
nrow(L7_records)/2


L7_records <- order_data(L7_records)
saveRDS(L7_records, paste0(Ldir, "L7_records.RDS"))

```


```{r}
L8_records <- readRDS(paste0(Ldir, "records_new.RDS"))
L8_records <- L8_records[!is.na(L8_records$download_available),]

L8_records <- L8_records[L8_records$level=="sr" | L8_records$level=="bt",] 

ggplot(L8_records, aes(x=start_time, y=cloudcov_land))+
  geom_point()+theme_minimal()+
  ggtitle(paste0("Landsat 8 data Nov 13 to Jan 14, n=", nrow(L8_records)/2))

# L8_records <- get_data(L8_records) # doesn't work

```


```{r}
L7_records <- readRDS(paste0(Ldir, "L7_records.RDS"))
L7_records <- L7_records[!is.na(L7_records$download_available),]
L7_records <- L7_records[L7_records$level=="sr" | L7_records$level=="bt",] 

ggplot(L7_records, aes(x=start_time, y=cloudcov_land))+
  geom_point()+theme_minimal()+
  ggtitle(paste0("Landsat 8 data Jan 2011 to March 13, n=", nrow(L7_records)/2))

# L7_records <- get_data(L7_records)

```

There is no Landsat 4/5 data available for that timeframe and AOI, I checked that via the EarthExplorer as well as it was unavailable at the time of checking via package.

## Downloading via espa.tools because get_data() doesn't work

# To do: run all downloads 

```{r,eval=F}
#install.packages("espa.tools", repos="http://R-Forge.R-project.org")

# divide up by level to sort into different folders
L8sr <- L8_records[L8_records$level == "sr",]
L8bt <- L8_records[L8_records$level == "bt",]

L8_rec_list <- list("sr"=L8sr, "bt"=L8bt)

L7sr <- L7_records[L7_records$level == "sr",]
L7bt <- L7_records[L7_records$level == "bt",]

L7_rec_list <- list("sr"=L7sr, "bt"=L7bt)

# # make dirs to sort files into
# dirstocreateL8 <- paste0(downloadpathL8, names(L8_rec_list), "/")
# for(i in seq(dirstocreateL8)){
#   dir.create(dirstocreateL8[i])
# }
# 
# dirstocreateL7 <- paste0(downloadpathL7, names(L7_rec_list), "/")
# for(i in seq(dirstocreateL7)){
#   dir.create(dirstocreateL7[i])
# }

# download function 
download_L <- function(downloadpath, recordslist){
                lapply(seq(2), function(l){ # for bt and for sr
                    print(paste0("starting with ", l, " (1= sr / 2=bt)"))
                    records <- recordslist[[l]]
                    dlp <- paste0(downloadpath, names(recordslist)[l], "/")
                    download_missed <- 0

                      # for(i in seq(records$order_id)){
                      for(i in c(12:nrow(records))){
                      print(paste0("starting with ", i))
                      print(c(records$summary[i]))
                      if(records$ordered[i]==TRUE){
                        
                        
                          d <- try(earthexplorer_download(usgs_eros_username=newUSGSUser,
                                             usgs_eros_password=newUSGSPW,
                                             output_folder = dlp, 
                                             verbose=T,
                                             ordernum = records$order_id[i],
                                             overwrite = T))
                          if(class(d) == "try-error"){
                            print(paste0("couldn't download ", i))
                            download_missed <- append(download_missed, i)
                            write.csv2(download_missed, paste0(downloadpath, "download_missed_l", l, ".csv"))
                          } else {
                            print(paste0("downloaded ", i))
                          }
                      } else {
                        print("scene was not ordered")
                      }
                      }

                })
              }



download_L(downloadpath=downloadpathL8, 
           recordslist=L8_rec_list)
download_L(downloadpath=downloadpathL7, 
           recordslist=L7_rec_list)

```

# To do: untar all with split versions

```{r, eval=F}
  
sdirs <- list.files(downloadpathL8, full.names = T)
sdirnam <- basename(sdirs)
for(d in seq(sdirs)){
    print(paste0("starting subdir ", sdirs[d]))
    sc <- list.files(sdirs[d], full.names =T)
    s <- list.files(sdirs[d], full.names = F)
    for(i in seq(sc)){
      print(paste0("starting untar ", i, "/", length(sc)))
      untar(sc[i],
            exdir=paste0("D:/Surface_Moisture/Landsat_data/L8/", 
                         sdirnam[d], "/",
                         tools::file_path_sans_ext(tools::file_path_sans_ext(s[i])),
                         "/"))
    }
}

sdirs <- list.files(downloadpathL7, full.names = T)
sdirnam <- basename(sdirs)
for(d in seq(sdirs)){
    print(paste0("starting subdir ", sdirs[d]))
    sc <- list.files(sdirs[d], full.names =T)
    s <- list.files(sdirs[d], full.names = F)
    for(i in seq(sc)){
      print(paste0("starting untar ", i, "/", length(sc)))
      untar(sc[i],
            exdir=paste0("D:/Surface_Moisture/Landsat_data/L7/", 
                         sdirnam[d], "/",
                         tools::file_path_sans_ext(tools::file_path_sans_ext(s[i])),
                         "/"))
    }
}


```

