---
title: "Look into satellite data"
author: "Maite"
date: "14 12 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=F, warning=F, echo=F, out.width = '150%')
```

## 
```{r}
library(raster)
library(sf)
library(here)
library(mapview)
library(stringr)
library(lubridate)
library(ggpubr)
library(tidyr)
library(ggplot2)
library(corrplot)
library(dplyr)

```


```{r}
Ldir <- "D:/Surface_Moisture/Landsat_data/"
L8path <- "D:/Surface_Moisture/Landsat_data/L8/"
L7path <- "D:/Surface_Moisture/Landsat_data/L7/"
aoi <- st_read(here("../aux_data/Levy_MDV_actually.shp"), quiet = TRUE)
antaproj <- "+proj=stere +lat_0=-90 +lat_ts=-71 +lon_0=0 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs"
aoianta <- st_transform(aoi, antaproj)
source(here("helpscripts/L_cloud_clean.R"))
source(here("helpscripts/readMeta.R"))
source(here("helpscripts/process_Landsat_BT.R"))
source(here("helpscripts/match_pix_points.R"))
template <- raster(here("../aux_data/template_new.tif"))
ds <- list.files(L8path, full.names=T)

```

# Retrieve logger locations
```{r,eval=F}
iButDir <- paste0(here("../calib_valid/iButtons/"))

gpkdir <- paste0(iButDir, "/gpkg_DDcorrection/")
f <- list.files(gpkdir,full.names=T)

iB_valley <- lapply(seq(f),function(i){
  st_read(f[[i]])
})


unique_iB_loc <- iB_valley[[3]]
unique_iB_loc <- unique_iB_loc[c("Site_Name", "geom")]

iB_cols <- lapply(c(1,2,5,6,7,8), function(v){
  iB_valley[[v]][c("Site_Name", "T", "Rh", "time", "geom")]
})

iB_all <- do.call("rbind", iB_cols)
iB_all$time <- with_tz(iB_all$time, "Pacific/Auckland") # needs to be converted every time because geopackage changes timezone
iB_all$iBut_GMT_time <- with_tz(iB_all$time, "GMT")
SoilS <- read_sf(paste0(here("../calib_valid/AWS/SoilStations/"), "/SoilS.gpkg"))

SoilS$Site_Name <- SoilS$station

SoilS_a <- SoilS[c("Site_Name", "geom")]

Lproj <- "+proj=stere +lat_0=-90 +lat_ts=-71 +lon_0=0 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs"
point_locs <- rbind(unique_iB_loc, SoilS_a)
point_locs_anta <- st_transform(point_locs, crs = crs(Lproj))

SoilS_data <- readRDS(here("../calib_valid/AWS/SoilStations/all_AWSS_all_years.rds"))
# match names to point_locs_anta to match after extraction 
SoilS_data$Site_Name <- NA
SoilS_data$Site_Name[SoilS_data$ID=="BP"] <- "Bull Pass "
SoilS_data$Site_Name[SoilS_data$ID=="BPE"] <- "Bull Pass East"
SoilS_data$Site_Name[SoilS_data$ID=="DJP"] <- "Don Juan Pond"
SoilS_data$Site_Name[SoilS_data$ID=="GH"] <- "Granite Harbour"
SoilS_data$Site_Name[SoilS_data$ID=="MP"] <- "Marble Point"
SoilS_data$Site_Name[SoilS_data$ID=="MtF"] <- "Mt. Fleming"
SoilS_data$Site_Name[SoilS_data$ID=="VV"] <- "Victoria Valley"

datehour <- paste0(SoilS_data$DATE, "_",
                          SoilS_data$HOUR/100)
SoilS_data$time <- as.POSIXlt(datehour, format="%Y-%m-%d_%H", 
                              tz="Pacific/Auckland")
SoilS_data$GMT_time <- with_tz(SoilS_data$time, tz="GMT")


AWSS_IDs <- c("Bull Pass ", "Bull Pass East", 
              "Don Juan Pond", "Granite Harbour",
              "Marble Point", "Mt. Fleming",
              "Victoria Valley")
iBut_IDs <- point_locs_anta$Site_Name[1:94]


```

# Data preparation: 
d=type of Landsat data (1=bt,2=sr)

## BT to LST 

```{r, eval=F}
# bt
d=1
fd <- list.files(ds[d],full.names=T)
lstpath <- paste0("D:/Surface_Moisture/Landsat_data/", "LST/")
if(!dir.exists(lstpath)){
  dir.create(lstpath)
}

# cloudclean, mask by aoi, convert to BTC and LST and write 
lapply(seq(fd), function(i){
  processLandsatBT(fd[[i]])
})
```


```{r, eval=F}
lstf <- list.files(fd, pattern="_LST.tif", full.names=T)

# make a list of LST files and acquisition dates 
LSTs <- lapply(seq(lstf), function(i){
    mtlFile <- list.files(dirname(lstf[i]), pattern="MTL.txt", full.names = T)
    metaData <- readMeta(mtlFile)
    Lsc_date<- metaData$ACQUISITION_DATE
    LST <- raster(lstf[i])
    return(list(LST, Lsc_date))
})

LSTs_date <- lapply(LSTs,"[[",2)
LSTs <- lapply(LSTs, "[[",1)

# match pixels to points 
L_LST_match <- lapply(seq(LSTs), function(i){
  print(paste0("starting with ", i))
  match_pix_points(sat_scene=LSTs[[i]], sattype="LST",
                    point_locs=point_locs_anta,
                    date = LSTs_date[[i]],
                    maxna=1, sat_min_tolerance = 45,
                    iB_min_tolerance = 5,
                    SoilS_min_tolerance = 30)
})

(matched_LST_AWSS <- sapply(L_LST_match, "[[", 4))
(matched_LST_iB <- sapply(L_LST_match, "[[", 5))

data.frame(LScenenum = seq(1:length(L_LST_match)),
           locinfo=sapply(L_LST_match, "[[", 3),
           matched_LST_AWSS, matched_LST_iB)

AWSS_LST_pix <- lapply(L_LST_match, "[[", 1)
iBut_LST_pix <- lapply(L_LST_match, "[[", 2)

```

```{r, eval=F}

AWSS_LST_pix <- AWSS_LST_pix[!is.na(AWSS_LST_pix)]
AWSS_LST_full <- do.call("rbind", AWSS_LST_pix)

iBut_LST_pix <- iBut_LST_pix[!is.na(iBut_LST_pix)]
iBut_LST_full <- do.call("rbind", iBut_LST_pix)

write.csv2(AWSS_LST_full, here("../calib_valid/pix_SM/AWSS_LST_10scenes.csv"))
write.csv2(iBut_LST_full, here("../calib_valid/pix_SM/iBut_LST_10scenes.csv"))

```


## SR
lsr_landsat_8_c1 
lsr_landsat_etm_c1
[Info from USGS on scaling](https://www.usgs.gov/faqs/how-do-i-use-scale-factor-landsat-level-2-science-products)

Scale factor for C1 surface reflectance is 0.0001, Fill value is -9999 and valid range 0 to 10000. 

Scale factor was applied, but valid range not applied

Landsat 8 acquisition time comes in GMT. 
iBut data was gathered in NZDT, but comes in the geopackage in CET/CEST...  ???


```{r, eval=F}
#sr
d=2 
fd <- list.files(ds[d],full.names=T)

# processing sr files 
# scc <- lapply(c(8:10), function(f){
scc <- lapply(seq(fd), function(f){
  print(paste("starting with", f))
  fls <- list.files(fd[f], pattern="sr_band..tif", full.names = T)
  pqa <- list.files(fd[f], pattern="pixel_qa.tif", full.names = T)
  mtlFile <- list.files(fd[f], pattern="MTL.txt", full.names = T)
  
  sn <- tools::file_path_sans_ext(basename(fls))
  
  s <- raster::stack(fls)
  qa <- raster::raster(pqa)
  
  metaData <- readMeta(mtlFile)
  Lsc_date<- metaData$ACQUISITION_DATE
  
  # cloud masking and scaling SR bands 2:7
  scc <- L_cloud_scale(bands=s[[2:7]], qualras=qa, sn=sn)
  
  bands <- scc[[1]]
  c <- scc[[2]]
  
  print(paste("writing cloud mask to ", 
              paste0(fd[f], "/", names(qa), "_mask.tif")))
  
  writeRaster(c, 
              paste0(fd[f], "/", names(qa), "_mask.tif"),
              overwrite=T)
  
  print(paste("writing bands to ", 
              paste0(fd[f], "/",names(s[[2:7]]), "_cloud_rm_scaled.tif")))
  
  writeRaster(bands, bylayer=T,
            paste0(fd[f], "/",names(s[[2:7]]), "_cloud_rm_scaled.tif"),
            overwrite=T)
  
  rm(bands, c, scc, s)
  gc()
  
}) 

 # checking whether it worked in all subfolders
rmf <- lapply(seq(fd), function(f){
  list.files(fd[f], pattern="_cloud_rm_scaled.tif", full.names = T)
})


unique(dirname(unlist(rmf)))
length(unique(basename(unlist(rmf))))/6

 # gathering sr files 
scc <- lapply(seq(fd), function(f){
  print(f)
  fls <- list.files(fd[f], pattern="_cloud_rm_scaled.tif", full.names = T)
  mtlFile <- list.files(fd[f], pattern="MTL.txt", full.names = T)
  
  s <- raster::stack(fls)

  metaData <- readMeta(mtlFile)
  Lsc_date<- metaData$ACQUISITION_DATE
  
  return(list(s, Lsc_date))
}) 


scc_date <- lapply(scc,"[[",2)
scc <- lapply(scc, "[[",1)

# match pixels to points 
L_sr_match <- lapply(seq(scc), function(i){
  match_pix_points(sat_scene=scc[[i]], sattype="sr",
                    point_locs=point_locs_anta,
                    date = scc_date[[i]],
                    maxna=5, sat_min_tolerance = 45,
                    iB_min_tolerance = 5,
                    SoilS_min_tolerance = 30)
})


(matched_sr_AWSS <- sapply(L_sr_match, "[[", 4))
(matched_sr_iB <- sapply(L_sr_match, "[[", 5))

data.frame(LScenenum = seq(1:length(L_sr_match)),
           locinfo=sapply(L_sr_match, "[[", 3),
           matched_sr_AWSS, matched_sr_iB)

AWSS_sr_pix <- lapply(L_sr_match, "[[", 1)
iBut_sr_pix <- lapply(L_sr_match, "[[", 2)


```


bands 2 (b), 3(g), 4(r), 5(NIR), 6 (SWIR1), 7 (SWIR2), 10(thermal)
```{r, eval=F}

AWSS_sr_pix <- AWSS_sr_pix[!is.na(AWSS_sr_pix)]
AWSS_sr_full <- do.call("rbind", AWSS_sr_pix)

iBut_sr_pix <- iBut_sr_pix[!is.na(iBut_sr_pix)]
iBut_sr_full <- do.call("rbind", iBut_sr_pix)

write.csv2(iBut_sr_full, here("../calib_valid/pix_SM/iBut_sr_10scenes.csv"))
write.csv2(AWSS_sr_full, here("../calib_valid/pix_SM/AWSS_sr_10scenes.csv"))

iBut_sr_full <- read.csv2(here("../calib_valid/pix_SM/iBut_sr_10scenes.csv"))
AWSS_sr_full <- read.csv2(here("../calib_valid/pix_SM/AWSS_sr_10scenes.csv"))

iBut_LST_full <- read.csv2(here("../calib_valid/pix_SM/iBut_LST_10scenes.csv"))
AWSS_LST_full <- read.csv2(here("../calib_valid/pix_SM/AWSS_LST_10scenes.csv"))

# put SR and LST data frames together 
AWSS <- dplyr::bind_rows(AWSS_sr_full, AWSS_LST_full)
iBut <- dplyr::bind_rows(iBut_sr_full, iBut_LST_full)

write.csv2(iBut, here("../calib_valid/pix_SM/iBut_10scenes.csv"))
write.csv2(AWSS, here("../calib_valid/pix_SM/AWSS_10scenes.csv"))
```


# Figures 

```{r}
iBut <- read.csv2(here("../calib_valid/pix_SM/iBut_10scenes.csv"))
AWSS <- read.csv2(here("../calib_valid/pix_SM/AWSS_10scenes.csv"))

iBut$Lsc_date <- as.POSIXct(iBut$Lsc_date)
iBut$iBut_GMT_time <- as.POSIXct(iBut$iBut_GMT_time)

AWSS$Lsc_date <- as.POSIXct(AWSS$Lsc_date)
AWSS$GMT_time <- as.POSIXct(AWSS$GMT_time)

```


# iButton - RH 
```{r}
# aggregate so that all bands are in one variable 

iBut[,grep("tiv", names(iBut))] <- NULL # remove time intervals

iBut_long <- gather(iBut, band, sr_LST, 
                    band2:band7, LST, factor_key=TRUE)
band2to7_LST_cols <- c("#0e20e6", "#277924", "#D82011", "#9aec98",
         "#Ec98e4", "#730668", "#eb911c")
# plot(1:length(band2to7_LST_cols), 1:length(band2to7_LST_cols), 
#      col = band2to7_LST_cols, cex = 3, pch=19)
SM_RH_col <- "#0bc9de"
manshapes <- c(19, 15, 17, 18, 1, 2, 5, 8, 13, 9, 6, 7, 3, 4, 0)
```


## iButton Surface Reflectance vs. RH all sites together
```{r}
scaleFactor <-  max(iBut_long$Rh, na.rm=T) / 1
iBut_SR_fullplot <- ggplot(iBut_long)+
  geom_point(aes(Lsc_date, sr_LST, shape=Site, color=band))+
  geom_point(aes(iBut_GMT_time, Rh/scaleFactor, shape=Site), cex=2, 
             color=SM_RH_col)+
  xlab("date")+
  scale_color_manual(values = band2to7_LST_cols)+
  scale_y_continuous(limits=c(0,1),
        # Features of the first axis
        name = "SR",
        # Add a second axis and specify its features
        sec.axis = sec_axis(~.*scaleFactor, name="RH (%)")
  ) +theme_minimal()+
  theme(axis.title.y.left=element_text(color="black",size=11, face="bold"),
        axis.text.y.left=element_text(color="black",size=11),
        axis.title.y.right=element_text(color=SM_RH_col,size=11, face="bold"),
        axis.text.y.right=element_text(color=SM_RH_col, size=11))+ 
  ggtitle("Surface Reflectance and RH", subtitle = "by logger site (shape) and satellite band (color)")
iBut_SR_fullplot
```

## iButton LST vs. RH all sites together 
```{r}
# scaleFactor = range of RH / range of LST 
scaleFactor <- abs(100/(range(iBut$LST, na.rm=T)[1]-range(iBut$LST, na.rm=T)[2]))

# sec_axis(~.*scaleFactor+addval) tilde stands for first axis, using min val on
# first axis (iBut$LST) and subtracting this value from 0 (min val in RH)
addval <- 0-min(iBut$LST, na.rm=T)*scaleFactor

# ggplot(iBut)+
#   # geom_point(aes(Lsc_date, LST, shape=Site), color=band2to7_LST_cols[7])+
#   geom_point(aes(iBut_GMT_time, Rh, shape=Site), cex=2, 
#              color=SM_RH_col)

# scaleFactor <-  max(iBut$Rh, na.rm=T) / max(iBut$LST, na.rm=T)
iBut_LST_fullplot <- ggplot(iBut)+
  geom_point(aes(Lsc_date, LST, shape=Site), color=band2to7_LST_cols[7])+
  geom_point(aes(iBut_GMT_time, (Rh-addval)/scaleFactor, shape=Site), cex=2, 
             color=SM_RH_col)+
  xlab("date")+
  scale_shape_manual(values = manshapes[1:length(unique(iBut$Site))]) +
  scale_color_manual(values = band2to7_LST_cols[7])+
  scale_y_continuous(
        # Features of the first axis
        name = "LST",
        # Add a second axis and specify its features
        sec.axis = sec_axis(~.*scaleFactor+addval, name="RH (%)")
  ) +theme_minimal()+
  theme(axis.title.y.left=element_text(color="black",size=11, face="bold"),
        axis.text.y.left=element_text(color="black",size=11),
        axis.title.y.right=element_text(color=SM_RH_col,size=11, face="bold"),
        axis.text.y.right=element_text(color=SM_RH_col, size=11))+ 
  ggtitle("LST and RH", subtitle = "by logger site (shape) and satellite band (color)")

iBut_LST_fullplot
```

## iButton SR and LST vs. RH figure per site
```{r, SR_RH}
iBut_split <- split(iBut_long, iBut_long$Site)

RH_SR_split_plot_function <- function(data, i){
    scaleFactor <-  max(data$Rh, na.rm=T) / 1
    ggplot(data)+
      geom_point(aes(Lsc_date, sr_LST, shape=Site, color=band))+
      geom_point(aes(iBut_GMT_time, Rh/scaleFactor, shape=Site), cex=2, 
                 color=SM_RH_col)+
      xlab("date")+
      scale_color_manual(values = band2to7_LST_cols)+
      scale_shape_manual(values = manshapes[i]) +
      scale_y_continuous(limits=c(0,1),
            # Features of the first axis
            name = "SR",
            # Add a second axis and specify its features
            sec.axis = sec_axis(~.*scaleFactor, name="RH (%)")
      ) +theme_minimal()+
      theme(axis.title.y.left=element_text(color="black",size=11, face="bold"),
            axis.text.y.left=element_text(color="black",size=11),
            axis.title.y.right=element_text(color=SM_RH_col,size=11, face="bold"),
            axis.text.y.right=element_text(color=SM_RH_col, size=11))+ 
      ggtitle("Surface Reflectance and RH", subtitle = "by logger site (shape) and satellite band (color)")
}

splitplots_sr <- lapply(seq(iBut_split), function(i){
    RH_SR_split_plot_function(iBut_split[[i]], i)
})
  
```


```{r, LST_RH}
iBut_LST_split <- split(iBut, iBut$Site)


RH_LST_split_plot_function <- function(data, i){
  
        # scaleFactor = range of RH / range of LST 
      scaleFactor <- abs(100/(range(data$LST, na.rm=T)[1]-
                                range(data$LST, na.rm=T)[2]))
      
      # sec_axis(~.*scaleFactor+addval) tilde stands for first axis, using min val on
      # first axis (iBut$LST) and subtracting this value from 0 (min val in RH)
      addval <- 0-min(data$LST, na.rm=T)*scaleFactor
      
      # scaleFactor <-  max(iBut$Rh, na.rm=T) / max(iBut$LST, na.rm=T)
      ggplot(data)+
        geom_point(aes(Lsc_date, LST, shape=Site), color=band2to7_LST_cols[7])+
        geom_point(aes(iBut_GMT_time, (Rh-addval)/scaleFactor, shape=Site), cex=2, 
                   color=SM_RH_col)+
        xlab("date")+
        scale_color_manual(values = band2to7_LST_cols[7])+
        scale_shape_manual(values = manshapes[i]) +
        scale_y_continuous(
              # Features of the first axis
              name = "LST",
              # Add a second axis and specify its features
              sec.axis = sec_axis(~.*scaleFactor+addval, name="RH (%)")
        ) +theme_minimal()+
        theme(axis.title.y.left=element_text(color="black",size=11, face="bold"),
              axis.text.y.left=element_text(color="black",size=11),
              axis.title.y.right=element_text(color=SM_RH_col,size=11, face="bold"),
              axis.text.y.right=element_text(color=SM_RH_col, size=11))+ 
        ggtitle("LST and RH", subtitle = "by logger site (shape) and satellite band (color)")
      

}

splitplots_LST <- lapply(seq(iBut_LST_split), function(i){
    RH_LST_split_plot_function(iBut_LST_split[[i]], i)
})
```


```{r, fig.height=3, fig.width=10, results='hide'}

lapply(seq(splitplots_sr), function(i){
    ggarrange(splitplots_sr[[i]], splitplots_LST[[i]])
})
          

```


# AWSS - SMC

```{r}
# aggregate so that all bands are in one variable 
AWSS[,grep("tiv", names(AWSS))] <- NULL

AWSS_long <- gather(AWSS, band, sr_LST, 
                    band2:band7, LST, factor_key=TRUE)
band2to7_LST_cols <- c("#2ea3c4", "#277924", "#D82011", "#9aec98",
         "#Ec98e4", "#730668", "#eb911c")

```

## AWSS Surface Reflectance vs. SMC all sites together
```{r}
scaleFactor <-  max(AWSS_long$SMC, na.rm=T) / 1
ggplot(AWSS_long)+
  geom_point(aes(Lsc_date, sr_LST, shape=Site, color=band))+
  geom_point(aes(GMT_time, SMC/scaleFactor, shape=Site), cex=2, 
             color=SM_RH_col)+
  xlab("date")+
  scale_color_manual(values = band2to7_LST_cols)+
  scale_y_continuous(limits=c(0,1),
        # Features of the first axis
        name = "SR",
        # Add a second axis and specify its features
        sec.axis = sec_axis(~.*scaleFactor, name="SMC (wfv)")
  ) +theme_minimal()+
  theme(axis.title.y.left=element_text(color="black",size=11, face="bold"),
        axis.text.y.left=element_text(color="black",size=11),
        axis.title.y.right=element_text(color=SM_RH_col,size=11, face="bold"),
        axis.text.y.right=element_text(color=SM_RH_col, size=11))+ 
  ggtitle("Surface Reflectance and SMC", subtitle = "by AWS site (shape) and satellite band (color)")

```

## AWSS LST vs. SMC all sites together
```{r}
# scaleFactor = max SMC / range of LST 
scaleFactor <- abs(max(AWSS$SMC)/(range(AWSS$LST, na.rm=T)[1]-range(AWSS$LST, na.rm=T)[2]))

# sec_axis(~.*scaleFactor+addval) tilde stands for first axis, using min val on
# first axis (iBut$LST) and subtracting this value from 0 (min val in SMC)
addval <- 0-min(AWSS$LST, na.rm=T)*scaleFactor
# 
# AWSS$GMT_time_d <- as.Date(AWSS$GMT_time)
# AWSS$Lsc_date_d <- as.Date(AWSS$Lsc_date)

mindate <- min(AWSS$Lsc_date)
maxdate <- max(AWSS$Lsc_date)
library(scales)
# scaleFactor <-  max(iBut$Rh, na.rm=T) / max(iBut$LST, na.rm=T)
AWSS_LST_fullplot <- ggplot(AWSS)+
  geom_point(aes(Lsc_date, LST, shape=Site), color=band2to7_LST_cols[7])+
  geom_point(aes(GMT_time, (SMC-addval)/scaleFactor, shape=Site), cex=2,
             color=SM_RH_col)+
  # scale_x_date(date_breaks="3 days"
  # #limits = as.Date("2013-10-27", "2013-11-10")
  # )+
  xlab("date")+
  scale_shape_manual(values = manshapes[1:length(unique(AWSS$Site))]) +
  scale_color_manual(values = band2to7_LST_cols[7])+
  # scale_x_date(limits = as.Date(c(mindate,maxdate))) +

  scale_y_continuous(
        # Features of the first axis
        name = "LST",
        # Add a second axis and specify its features
        sec.axis = sec_axis(~.*scaleFactor+addval, name="SMC (wfv)")
  ) +theme_minimal()+
  theme(axis.title.y.left=element_text(color="black",size=11, face="bold"),
        axis.text.y.left=element_text(color="black",size=11),
        axis.title.y.right=element_text(color=SM_RH_col,size=11, face="bold"),
        axis.text.y.right=element_text(color=SM_RH_col, size=11),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+ 
  ggtitle("LST and SMC", subtitle = "by logger site (shape) and satellite band (color)")

AWSS_LST_fullplot
```


## AWSS SR and LST vs. SMC figure per site 
```{r, SR_SMC}
AWSS_split <- split(AWSS_long, AWSS_long$Site)

SR_SMC_split_plot_function <- function(data, i){
    scaleFactor <-  max(data$SMC, na.rm=T) / 1
    ggplot(data)+
      geom_point(aes(Lsc_date, sr_LST, shape=Site, color=band))+
      geom_point(aes(GMT_time, SMC/scaleFactor, shape=Site), cex=2, 
                 color=SM_RH_col)+
      xlab("date")+
      scale_color_manual(values = band2to7_LST_cols)+
      scale_shape_manual(values = manshapes[i]) +
      scale_y_continuous(limits=c(0,1),
            # Features of the first axis
            name = "SR",
            # Add a second axis and specify its features
            sec.axis = sec_axis(~.*scaleFactor, name="SMC (wfv)")
      ) +theme_minimal()+
      theme(axis.title.y.left=element_text(color="black",size=11, face="bold"),
            axis.text.y.left=element_text(color="black",size=11),
            axis.title.y.right=element_text(color=SM_RH_col,size=11, face="bold"),
            axis.text.y.right=element_text(color=SM_RH_col, size=11))+ 
      ggtitle("Surface Reflectance and SMC", subtitle = "by AWSS site (shape) and satellite band (color)")
}

splitplots_AWSS_SR <- lapply(seq(AWSS_split), function(i){
    SR_SMC_split_plot_function(AWSS_split[[i]], i)
})

```

```{r, LST_SMC}
SMC_LST_split_plot_function <- function(data, i){
  
      data <- data[data$band=="LST",]
  
      # scaleFactor = max SMC / range of LST 
      scaleFactor <- abs(max(data$SMC)/
                           (range(data$sr_LST, na.rm=T)[1]-range(data$sr_LST, na.rm=T)[2]))
      
      # sec_axis(~.*scaleFactor+addval) tilde stands for first axis, using min val on
      # first axis (iBut$LST) and subtracting this value from 0 (min val in SMC)
      addval <- 0-min(data$sr_LST, na.rm=T)*scaleFactor

      
      # scaleFactor <-  max(iBut$Rh, na.rm=T) / max(iBut$LST, na.rm=T)
      ggplot(data)+
        geom_point(aes(Lsc_date, sr_LST, shape=Site), color=band2to7_LST_cols[7])+
        geom_point(aes(GMT_time, (SMC-addval)/scaleFactor, shape=Site), cex=2,
             color=SM_RH_col)+
        xlab("date")+
        scale_color_manual(values = band2to7_LST_cols[7])+
        scale_shape_manual(values = manshapes[i]) +
        scale_y_continuous(
              # Features of the first axis
              name = "LST",
              # Add a second axis and specify its features
              sec.axis = sec_axis(~.*scaleFactor+addval, name="SMC (wfv)")
        ) +theme_minimal()+
        theme(axis.title.y.left=element_text(color="black",size=11, face="bold"),
              axis.text.y.left=element_text(color="black",size=11),
              axis.title.y.right=element_text(color=SM_RH_col,size=11, face="bold"),
              axis.text.y.right=element_text(color=SM_RH_col, size=11))+ 
        ggtitle("LST and RH", subtitle = "by logger site (shape) and satellite band (color)")
      

}

splitplots_AWSS_LST <- lapply(seq(AWSS_split), function(i){
    SMC_LST_split_plot_function(data=AWSS_split[[i]], i)
})

```


```{r, fig.height=3, fig.width=10, results='hide'}

lapply(seq(splitplots_AWSS_SR), function(i){
    ggarrange(splitplots_AWSS_SR[[i]], splitplots_AWSS_LST[[i]])
})
          

```


There might be something about the relation between reflectance in the visible light and SWIR reflectance.. it seems like always when the moisture (RH or SMC) is relatively high, SWIR remains low, while VIS is relatively high.. When VIS reflectance is pretty high, SMC is low. 
This is visible also in the following corrplot: rgb (bands 4,3,2) and nir (5) all correlate with SMC around r=-0.35 to r=-0.4, while SWIR bands 6 and 7 are positively correlated with r=0.77 and 0.83:
For SMC it seems useful to try a VIS/SWIR index. SWIR band 6 seems best for exploiting the relation to VIS (measured reflectance was usually a bit lower than 7), especially with the green band 3, there is much negative correlation (-0.75). 

# Correlations AWSS and SMC

```{r, out.width = '150%'}
make_corrplot <- function(data, type, LST_or_sr){
  
  data$ind <- (data$band6-data$band3)/
    (data$band6+data$band3)
  data$diff2 <- (data$band6-data$band2)
  data$diff3 <- (data$band6-data$band3)
  data$diff4 <- (data$band6-data$band4)
  data$diff5 <- (data$band6-data$band5)
  
  b2pos <- which(names(data)=="band2")
  b7pos <- which(names(data)=="band7")
  indpos <- which(names(data)=="ind")
  diff5pos <- which(names(data)=="diff5")
  
  if(type=="iBut"){
    moist <- "Rh"
  } else {
    moist <- "SMC"
  }
  
  if(LST_or_sr=="sr"){
      cormatsr <- cor(data[,c(b2pos:b7pos,indpos:diff5pos,
                          which(names(data)==moist))],
                   use="pairwise.complete.obs")
     corrplot(cormatsr, method = "number", 
                       type = "lower", number.cex=0.8,
           tl.col = "black",insig="p-value")
  } else {
      cormatLST <- cor(data[,c(which(names(data)=="LST"),
                          which(names(data)==moist))],
                   use="pairwise.complete.obs")

      corrplot(cormatLST, method = "number", 
                        type = "lower", number.cex=1.3,
           tl.col = "black",insig="p-value")
  }
}

make_corrplot(data=AWSS, type="AWSS", LST_or_sr="sr")
```


```{r, fig.align="center", out.width = '50%'}
make_corrplot(data=AWSS, type="AWSS",LST_or_sr="LST")
```


For RH and the iButtons, the relation seems to be the opposite, bands 2:5 are quite highly positively correlated with Rh, the SWIR bands not really, though. 

# Correlations iButtons and RH

```{r, out.width = '150%'}
make_corrplot(data=iBut, type="iBut", LST_or_sr="sr")
```


```{r, fig.align="center", out.width = '50%'}
make_corrplot(data=iBut, type="iBut", LST_or_sr="LST")
```

## Synopsis
### SMC and data

* strong positive correlation SMC with SWIR
  * r=0.93 for SMC and SWIR band 7
  * r=0.77 for SMC and SWIR band 6
* medium negative correlation SMC with VNIR
* best difference index NIR and SWIR 6 (there is no difference when using band 7)

### RH and iButtons
* strong positive correlation with VIS bands 
* no good correlation with SWIR on its own
* strong negative correlation with difference band 6-VNIR
  * r=-0.69 for RH and diff SWIR band 6 - NIR
  * r=-0.68 for RH and diff SWIR band 6 - R or G or B 
* differences: there is no change when using band 7 instead of band 6
