---
title: "Look into other predictors"
author: "Maite"
date: "24 1 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=F, warning=F, echo=F, out.width = '150%')
```

```{r}
library(raster)
library(mapview)
library(here)
library(sf)
```

## read in predictor datasets

* DEM 
* rock outcrop
* slope
* soil type 
## to do: 
* once it is clear which dates are relevant: get wind speed and direction 
* terrain-derived data 
  * TWI
  * regional microclimate (small depressions, polygon cracks etc)
* distance to lakes and runoff channels
* pH // EC: Pierre
* permafrost data: Bockheim


```{r, eval=F}
# predpath <- here("../predictors/")
# dem <- raster(paste0(predpath, "/DEM_8m_MDV_filled_aoi.tif"))
# full_grids <- readRDS(paste0(predpath, "/full_size_grids_all_layers.RDS"))
# rock_out <- full_grids$landcoverres
# rock_out_8 <- resample(rock_out, dem)
# writeRaster(rock_out_8, paste0(predpath, "/rock_outcrop_8m.grd"))
# dem[rock_out_8 > 0.96] <- NA #0.97 = snow and ice, other values because of resampling
# writeRaster(dem, paste0(predpath, "/DEM_8m_MDV_filled_aoi_no_snow.grd"), overwrite=T)
# TWI <- raster(paste0(predpath, "/TWI_RSlope.tif"))
# TWI_res <- resample(TWI, dem)
# writeRaster(TWI_res, paste0(predpath, "TWI_resampled_dem.grd"))




```

Get logger point data 
```{r}
AWSS_loc <- read_sf(here("../calib_valid/AWS/SoilStations/SoilS.gpkg"))
iBut_loc <- read_sf(here("../calib_valid/iButtons/gpkg_DDcorrection/iBut_dec.gpkg"))
```


Crop all predictors to rock outcrop settings

```{r}
predpath <- here("../predictors/")
dem_full <- raster(paste0(predpath, "/DEM_8m_MDV_filled_aoi.tif"))
dem_nosnow <- raster(paste0(predpath, "/DEM_8m_MDV_filled_aoi_no_snow.grd"))
rock_out_8 <- raster(paste0(predpath, "/rock_outcrop_8m.grd"))
full_grids <- readRDS(paste0(predpath, "/full_size_grids_all_layers.RDS"))
slope <- full_grids$slope
rock_out <- full_grids$landcoverres
soil_type <- full_grids$soilraster

slope[rock_out > 0.96] <- NA
soil_type[rock_out > 0.96] <- NA

TWI <- raster(paste0(predpath, "/TWI_resampled_dem.grd"))

```

## Lakes and streams 


```{r, eval=F}
# For streams using https://data.bas.ac.uk/items/6a4d6fa2-d713-45ab-a30a-7130795e29d2/ Version 7.3
# Released: 2020-11-05T09:05:13 from the Antarctic Digital Database. 
# 'Gerrish, L., Fretwell, P., & Cooper, P. (2020). Antarctic streams dataset (7.3) [Data set]. UK Polar Data Centre, Natural Environment Research Council, UK Research & Innovation. https://doi.org/10.5285/6a4d6fa2-d713-45ab-a30a-7130795e29d2' 
# 
# High resolution Antarctic lakes Version: 7.3, Released:2020-11-05T09:03:57
# 'Gerrish, L., Fretwell, P., & Cooper, P. (2020). High resolution Antarctic lakes dataset (7.3) [Data set]. UK Polar Data Centre, Natural Environment Research Council, UK Research & Innovation. https://doi.org/10.5285/6a27ab9e-1258-49b1-bd2c-fcbed310ab45'.

lakes <- read_sf(here("../predictors/lakes_streams/add_lakes_high_res_polygon_v7.3.gpkg"))
streams <- read_sf(here("../predictors/lakes_streams/add_streams_v7.3.gpkg"))

```

Streams and lakes vectors from the Antarctic Digital Database (referenced in chunk above) don't really look that great... Got some Sentinel 2 scenes for detecting lakes and streams by myself: 

```{r}
lakes <- read_sf(here("../predictors/lakes.shp"))
streams <- read_sf(here("../predictors/streams.shp"))

iBut_loc_anta <- st_transform(iBut_loc, crs=crs(lakes))
AWSS_loc_anta <- st_transform(AWSS_loc, crs=crs(lakes))

iBalldist <-  st_distance(iBut_loc_anta, streams, by_element=F)
AWalldist <- st_distance(AWSS_loc_anta, streams, by_element=F)

# get minimum per row (feature in logger data)
iBut_loc_anta$dist_stream <- apply(iBalldist,1,min)
AWSS_loc_anta$dist_stream <- apply(AWalldist,1,min)


iBalldist_l <-  st_distance(iBut_loc_anta, lakes, by_element=F)
AWalldist_l <- st_distance(AWSS_loc_anta, lakes, by_element=F)

iBut_loc_anta$dist_lake <- apply(iBalldist,1,min)
AWSS_loc_anta$dist_lake <- apply(AWalldist,1,min)


```


```{r}
mapview(dem_nosnow, na.color="#00000000")+
  mapview(slope, na.color="#00000000")+
  mapview(soil_type, na.color="#00000000")+
  mapview(lakes, color="blue", col.regions="blue")+
  mapview(streams, color="blue")+
  mapview(AWSS_loc_anta, zcol="dist_stream")+
  mapview(iBut_loc_anta, cex=4,
          zcol="dist_stream")
```

## Characterize iButton locations by predictors
```{r}

terrainstack <- stack(dem_full, TWI)
stack30 <- stack(full_grids$slope, full_grids$aspect, full_grids$soilraster)
 
tsiB <- raster::extract(terrainstack, iBut_loc_anta)
tsAW <- extract(terrainstack, AWSS_loc_anta)


```

## Characterize AWSS locations by predictors
```{r}

```

