---
title: "AWS_Met_Stations"
author: "Maite"
date: "29 10 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r, echo=F, message=F, warning=F}
library(sf)
library(mapview)
library(ggplot2)
library(padr)
library(gridExtra)
library(grid)
library(colorspace)
library(here)
library(dplyr)
library(lubridate)
```


### Met Stations

Unfortunately, all of them can't be used, because Relative Humidity is measured at 3m height. 

Dataset can be downloaded here: [Met Data](https://mcm.lternet.edu/meteorology-data-sets)


```{r, echo=F, message=F, warning=F}
awsDir <- here("../calib_valid/AWS/")

ml <- read.csv(list.files(awsDir, pattern="METLOCS",full.names=T), sep=",")

ml <- ml[complete.cases(ml$LONGITUDE),]
ml <- ml[grepl("Met Station", ml$LOCATION_NAME),]


aoi <- st_read(here("../aux_data/Levy_MDV_actually.shp"))
dem <- raster::raster(here("../predictors/DEM_8m_MDV_filled_aoi.tif"))


# ml$running <- 0
# ml$YEAR_REMOVED[is.na(ml$YEAR_REMOVED)] <- 0
# ml$YEAR_REMOVED-ml$YEAR_INSTALLED
```


**Promising stations on soil (12) ** 

Mt. Fleming Met Station (AirT 2m and 1.3m aboveground, RH probably same heights)

Beacon Valley Met Station (soil temperature 0cm RH at 3m)

Friis Hills Met Station (AirT 3m above ground, RH at 3m)

Bonney Riegel Sensit/Theta Soil Met Station (= Snowfence station. no RH)

Upper Howard Met Station (RH at 3m)

* Garwood Valley Met Station (RH and soiltemperature 0m present, not sure *whether RH is measured in soil or at 3m height?*)

Garwood Ice Cliff Met Station (RH at 3m, no soil temperature)

* Miers Valley Met Station (there is soil temperature 0m, not sure *whether RH is measured in soil or at 3m height?* )

F6 Sensit/Theta Soil Met Station (no RH)

Lake Brownworth Met Station (soil temperature 0m, RH at 3m)

Lake Vanda Met Station (soil temperature 0m, RH at 3m)

* Explorer's Cove Met Station (there is soil temperature 0m, not sure *whether RH is measured in soil or at 3m height?* )

**Stations at the edge of lakes (3)** 

Lake Bonney Met Station (soil temperature 0m, RH at 3m)

Lake Vida Met Station (soil temperature 0m, RH at 3m)

Lake Fryxell Met Station (there is soil temperature 0m, not sure *whether RH is measured in soil or at 3m height?* )

**Stations on glaciers / lakes - not to use (6)**

* Taylor Glacier Met Station

* Howard Glacier Met Station 

* Lake Fryxell Center Ice Met Station

* Commonwealth Glacier Met Station

* Canada Glacier Met Station 

* Lake Hoare Met Station


The following map shows the research area and the AWS to be used in yellow, those excluded in red and those I'm not sure about in black. 

```{r, echo=F, message=F, warning=F}
ml_points <- st_as_sf(ml, coords = c("LONGITUDE", "LATITUDE"), 
         crs = 4326)

use <- c("use", "use", "not", 
         "not", "use", "use", 
         "use", "use", "use",
         "not", "maybe", "use", 
         "not", "maybe", "not", 
         "use", "maybe", "use",
         "use", "not", "use")
names(use) <- ml_points$LOCATION_NAME
ml_points$use <- as.factor(use)

mapview::mapview(ml_points, label="LOCATION_NAME",
                 col.regions = mapviewGetOption("raster.palette"),
                 zcol="use", map.types="Esri.WorldImagery")+
  mapview::mapview(aoi, map.types = c("Esri.WorldImagery"),
                 alpha.regions = 0,
                 color = 'chartreuse3',
                 lwd=3)+
  mapview::mapview(dem,na.color="#00000000", map.types = "Esri.WorldImagery")
```



```{r, echo=F, message=F, warning=F}
lf_rh <- list.files(awsDir, full.names=T, pattern="frlm_rh")
lf_st <- list.files(awsDir, full.names=T, pattern="frlm_soilt")

f_rh <- read.csv(lf_rh)
head(f_rh, n=3)
# str(f_rh)
# unique(f_rh$RH_COMMENTS)
f_rh$datetime <- as.POSIXct(f_rh$DATE_TIME, format="%d/%m/%Y %H:%M")

f_st <- read.csv(lf_st)
head(f_st, n=3)
# str(f_st)
# unique(f_st$SOILT_COMMENTS)
f_st$datetime <- as.POSIXct(f_st$DATE_TIME, format="%d/%m/%Y %H:%M")
# unique(year(f_st$datetime))
```

Years that are covered in this dataset are: 
```{r, echo=F, message=F, warning=F}
hist(year(f_rh$datetime), main="RH Observations per year", 
     xlab="Year", ylab="n observations")
hist(year(f_st$datetime), main="Soil Temperature Observations per year", 
     xlab="Year", ylab="n observations")
```



Let's for now take a look at the year 2013 (we've got data from Taylor valley 
for that season) and focus on Relative Humidity and 
the soil temperature at surface level. We're excluding timestamps where surface 
soil temperature is below -1°C. When soil and surface water freezes no reliable 
relative humidity information can be measured any more. 
```{r, echo=F, message=F, warning=F}
f_rh13 <- f_rh[year(f_rh$datetime)==2013,]
f_st13 <- f_st[year(f_st$datetime)==2013,]

# filling the gaps in incomplete datetime variables so that no 
#data are actually no data in the plot
f_rh13 <- pad(f_rh13) 
f_st13 <- pad(f_st13)

# which timestamps are useable? 
f_rh13$useable <- 1
f_rh13$useable[f_st13$SOILT0CM < (-1)] <- 0
f_rh13$useable <- factor(f_rh13$useable)

f_st13$useable <- 1
f_st13$useable[f_st13$SOILT0CM < (-1)] <- 0
f_st13$useable <- factor(f_st13$useable)

prh <- ggplot(f_rh13, aes(datetime, RH, color=useable))+
  geom_point(size=0.5)+
  scale_color_manual(values=c("#e1e2e6","#0d0d0c"))+
  ggtitle("Relative Humidity")+
  theme_minimal()
pst <- ggplot(f_st13, aes(datetime, SOILT0CM, color=useable))+
  geom_point(size=0.5)+
  geom_hline(yintercept = -1, lty="dashed")+
  scale_color_manual(values=c("#e1e2e6","#0d0d0c"))+
  ggtitle("Soil temperature at surface level")+
  theme_minimal()

gridExtra::grid.arrange(prh,pst)
```
Weird pattern! Why only data for half the month but frosting and defrosting during
the whole year?! 

Zoom in on one month
```{r, echo=F, message=F, warning=F}
f_rh13_jan <- f_rh13[month(f_rh13$datetime)==3,]
f_st13_jan <- f_st13[month(f_st13$datetime)==3,]

prh <- ggplot(f_rh13_jan, aes(datetime, RH, color=useable))+
  geom_point(size=0.5)+
  geom_line()+
  scale_color_manual(values=c("#e1e2e6","#0d0d0c"))+
  ggtitle("Relative Humidity")+
  theme_minimal()
pst <- ggplot(f_st13_jan, aes(datetime, SOILT0CM, color=useable))+
  geom_point(size=0.5)+
  geom_line()+
  geom_hline(yintercept = -1, lty="dashed")+
  scale_color_manual(values=c("#e1e2e6","#0d0d0c"))+
  ggtitle("Soil temperature at surface level")+
  theme_minimal()

gridExtra::grid.arrange(prh,pst)
```


