---
title: "Prepare AWS SoilStations dataset"
author: "Maite"
date: "2 12 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=F, warning=FALSE, message=FALSE}
library(sf)
library(mapview)
library(ggplot2)
library(padr)
library(gridExtra)
library(grid)
library(colorspace)
library(here)
library(dplyr)
library(lubridate)
library(readxl)
library(raster)
```

### Read Soil Station data

Files were read and non-empty columns retained. A column name from the 
first three lines of the excel files was created. Finally, the so created
column names were matched to the pattern for the following variables of 
interest: 


ID, 
Year, 
Hour, 
Date
Air Temp (°C, 1.6m)
Solar Radiation (W/m²), 3m
Wind Speed (m/s), 3m
Wind Direction (azimuth), 3m

Soil Temp (°C) Surface under white rock
Soil Temp (°C) Surface under black rock

Soil RH (%) 5cm
Soil RH (%) 15cm

MRC Soil Temp °C 0 in 
MRC Soil Temp °C 3 in 7,6cm

Soil 2cm: 
ER	unitless	Real dielectric constant
EI	unitless	Imaginary dielectric constant
Temp	°C	Temperature
ER_COR	unitless	Temperature corrected ER
EI_COR	unitless	Temperature corrected EI
WATER	wfv	Soil moisture
SALINITY	g NaCl/L	Soil salinity
SOIL_COND	S/m	Soil conductivity
SOIL_COND_COR	S/m	Temp corrected soil conductivity
WATER_CON_COR	S/m	Temp corrected conductivity of the water in the soil


```{r, warning=FALSE, message=FALSE}
sdirs <- list.dirs(here("../calib_valid/AWS/SoilStations/"))[-1]

# PATTERNS FOR SUBSETTING THE DATASETS 
# BULL PASS     
pat_BP <- c("ID","YEAR_NA_NA", "HOUR_NA_NA", "DATE_NA_NA",
         "Air_Temp_deg_C", "Solar_Radiation_W/m2",
         "Wind_Speed_m/s", "Wind_Direction_deg_True_N",
         "Soil_Temp_2_cm_White_deg_C", "Soil_Temp_2_cm_Black_deg_C",
         "MRC_0_cm_deg_C", "MRC_7.5_cm_deg_C", "SOIL_RH_%", "SOIL_RH_%", 
         "ER_2_cm,_1_NA", "EI_2_cm,_1_NA","TEMP_2_cm,_1_deg_C",
         "ER_COR_2_cm,_1_NA","EI_COR_2_cm,_1_NA","WATER_2_cm,_1_wfv" ,
         "SALINITY_2_cm,_1_g_NaCl/L",   "SOIL_COND_2_cm,_1_S/m","SOIL_COND_COR_2_cm,_1_S/m" ,
         "WATER_CON_COR_2_cm,_1_S/m")

# BULL PASS EAST 
# don't exist: "Soil_Temp_2_cm_White_deg_C", "Soil_Temp_2_cm_Black_deg_C", 
# MRC_5.5_cm_deg_C instead of 7.5cm 
# instead of "SOIL_RH_%" there is "Soil_RH_2cm_%" 
# Bull Pass East from 2019 is mostly empty 
pat_BPE <- c("ID","YEAR_NA_NA", "HOUR_NA_NA", "DATE_NA_NA",
         "Air_Temp_deg_C", "Solar_Radiation_W/m2","Relative_Humidity_%",
         "Wind_Speed_m/s", "Wind_Direction_deg_True_N",
         "MRC_0_cm_deg_C", "MRC_5_cm_deg_C", 
         "Soil_RH_2_cm_%","Soil_Temp_2_cm_deg_C",
         "ER_2_cm,_1_NA", "EI_2_cm,_1_NA","TEMP_2_cm,_1_deg_C",
         "ER_COR_2_cm,_1_NA","EI_COR_2_cm,_1_NA","WATER_2_cm,_1_wfv" ,
         "SALINITY_2_cm,_1_g_NaCl/L",   "SOIL_COND_2_cm,_1_S/m","SOIL_COND_COR_2_cm,_1_S/m" ,
         "WATER_CON_COR_2_cm,_1_S/m")

# DON JUAN POND
# don't exist: "Soil_Temp_2_cm_White_deg_C", "Soil_Temp_2_cm_Black_deg_C", 
# MRC_5_cm_deg_C instead of 7.5cm 
# instead of "SOIL_RH_%", there's "Soil_RH_2_cm_%" 
# 2019: data missing 
pat_DJP <- c("ID","YEAR_NA_NA", "HOUR_NA_NA", "DATE_NA_NA",
         "Air_Temp_deg_C", "Solar_Radiation_W/m2",
         "Wind_Speed_m/s", "Wind_Direction_deg_True_N",
         "MRC_0_cm_deg_C", "MRC_5_cm_deg_C", "Soil_RH_2_cm_%", 
         "ER_2_cm,_1_NA", "EI_2_cm,_1_NA","TEMP_2_cm,_1_deg_C",
         "ER_COR_2_cm,_1_NA","EI_COR_2_cm,_1_NA","WATER_2_cm,_1_wfv" ,
         "SALINITY_2_cm,_1_g_NaCl/L",   "SOIL_COND_2_cm,_1_S/m","SOIL_COND_COR_2_cm,_1_S/m" ,
         "WATER_CON_COR_2_cm,_1_S/m")
# GRANITE HARBOUR
# don't exist: "Soil_Temp_2_cm_White_deg_C", "Soil_Temp_2_cm_Black_deg_C", 
# Air_Temp,_1_deg_C 
# MRC MRC_5.5_cm_deg_C instead of 7.5 or 5
# No soil RH
# ER,, EI, etc 2.5cm instead of 2cm 
pat_GH <- c("ID","YEAR_NA_NA", "HOUR_NA_NA", "DATE_NA_NA",
         "Air_Temp,_1_deg_C", "Solar_Radiation_w/m2",
         "Wind_Speed_m/s", "Wind_Direction_deg_True_N",
         "MRC_0_cm_deg_C", "MRC_5.5_cm_deg_C", 
         "ER_2.5_cm_NA","EI_2.5_cm_NA","TEMP_2.5_cm_deg_C","ER_COR_2.5_cm_NA",
         "EI_COR_2.5_cm_NA","WATER_2.5_cm_wfv","SALINITY_2.5_cm_g_NaCl/L",
         "SOIL_COND_2.5_cm_S/m","SOIL_COND_COR_2.5_cm_S/m","WATER_CON_COR_2.5_cm_S/m" )

# MARBLE POINT
# don't exist: "Soil_Temp_2_cm_White_deg_C", "Soil_Temp_2_cm_Black_deg_C", 
# "SOIL_RH_%" doesn't exist, instead: 
pat_MP <-  c("ID","YEAR_NA_NA", "HOUR_NA_NA", "DATE_NA_NA",
         "Air_Temp_deg_C", "Solar_Radiation_W/m2",
         "Wind_Speed_m/s", "Wind_Direction_deg_True_N",
         "MRC_0_cm_deg_C", "MRC_7.5_cm_deg_C", 
         "ER_2_cm,_1_NA", "EI_2_cm,_1_NA","TEMP_2_cm,_1_deg_C",
         "ER_COR_2_cm,_1_NA","EI_COR_2_cm,_1_NA","WATER_2_cm,_1_wfv" ,
         "SALINITY_2_cm,_1_g_NaCl/L",   "SOIL_COND_2_cm,_1_S/m","SOIL_COND_COR_2_cm,_1_S/m" ,
         "WATER_CON_COR_2_cm,_1_S/m")

# MOUNT FLEMMING 
# don't exist: "Soil_Temp_2_cm_White_deg_C", "Soil_Temp_2_cm_Black_deg_C"
# no MRC, instead  "Soil_Temp_Poly2cm_deg_C"     "Soil_Temp_Poly7.5cm_deg_C"
# no SOIL_RH_%
pat_MtF <- c("ID","YEAR_NA_NA", "HOUR_NA_NA", "DATE_NA_NA",
         "Air_Temp_deg_C", "Solar_Radiation_W/m2",
         "Wind_Speed_m/s", "Wind_Direction_deg_True_N",
          "Soil_Temp_Poly2cm_deg_C","Soil_Temp_Poly7.5cm_deg_C",
         "ER_7.5_cm,_1_NA","EI_7.5_cm,_1_NA","TEMP_7.5_cm,_1_deg_C",
         "ER_COR_7.5_cm,_1_NA","EI_COR_7.5_cm,_1_NA", "WATER_7.5_cm,_1_wfv",   
         "SALINITY_7.5_cm,_1_g_NaCl/L", "SOIL_COND_7.5_cm,_1_S/m",
         "SOIL_COND_COR_7.5_cm,_1_S/m" ,"WATER_CON_COR_7.5_cm,_1_S/m")

# don't exist: "Soil_Temp_2_cm_White_deg_C", "Soil_Temp_2_cm_Black_deg_C"
# no soil RH
# 
pat_VV <- c("ID","YEAR_NA_NA", "HOUR_NA_NA", "DATE_NA_NA",
         "Air_Temp_deg_C", "Solar_Radiation_w/m2",
         "Wind_Speed_m/s", "Wind_Direction_deg_True_N",
         "MRC_0_cm_deg_C", "MRC_7.5_cm_deg_C", 
         "ER_3_cm,_2_NA","EI_3_cm,_2_NA","TEMP_3_cm,_2_deg_C","ER_COR_3_cm,_2_NA",      
         "EI_COR_3_cm,_2_NA","WATER_3_cm,_2_wfv","SALINITY_3_cm,_2_g_NaCl/L", 
         "SOIL_COND_3_cm,_2_S/m","SOIL_COND_COR_3_cm,_2_S/m" , "WATER_CON_COR_3_cm,_2_S/m")


pat_list <- list(pat_BP, pat_BPE, pat_DJP, pat_GH, pat_MP, pat_MtF, pat_VV)
shortID <- list("BP", "BPE", "DJP", "GH", "MP", "MtF", "VV")

```

### Just for testing the patterns
```{r, warning=FALSE, message=FALSE, eval=F}
checktabs <- lapply(seq(sdirs), function(d){ # for all different stations
  print(paste0("d= ",d))
  f <- list.files(sdirs[d], full.names=T,pattern=".xls")
  stat_year <- lapply(seq(length(f)), function(y){ # for all years available for this station
  # loop through all of the files
      print(y)
      names_df <- read_excel(f[y],
           col_names = F)
      # get rid of empty columns
      any(is.na(names_df[5,]))
      
      isnadf <- is.na(names_df[200:216,])
      full_cols <- which(colSums(isnadf)!=16)
      
      pastenams <- apply(names_df[1:3,], 2, paste , collapse = "_" )
      pastenams <- gsub(" ", "_", pastenams)
      pastenams <- pastenams[full_cols]
      
      tab <- read_excel(f[y], .name_repair = "unique",
                 skip=3)
      
      isnatabdf <- is.na(tab[200:216,])
      full_cols_tab <- which(colSums(isnatabdf)!=16)
      tab <- tab[,full_cols_tab]
      ncol(tab)
      length(pastenams)
      
      names(tab) <- pastenams
      
      names(tab)[1] <- "ID"
      tab$ID <- shortID[[d]]
      
      return(tab[1:10,])
  })
})

names(checktabs) <- basename(sdirs)[1:length(checktabs)]
# saveRDS(checktabs, here("../calib_valid/AWS/SoilStations/checktabs.rds"))
```

Finding out what's going on: 
```{r, warning=FALSE, message=FALSE, eval=F}

# which variables in pat are in checktab names? 
check_varsel <- lapply(seq(checktabs), function(d){ # for all different stations
  sapply(seq(checktabs[[d]]), function(y){ # for all years available for this station
    sel <- which(names(checktabs[[d]][[y]]) %in% pat_list[[d]])
    length(sel)
    return(length(sel))
  })
})

names(check_varsel) <- basename(sdirs)[1:length(checktabs)]
check_varsel

# different naming for different stations? 

'%notin%' <- Negate('%in%')

# which columns could not be matched? 
pat_not_matched <- lapply(seq(checktabs), function(d){ # for all different stations
  s <- lapply(seq(checktabs[[d]]), function(y){ # for all years available for this station
    not_matched <- which(pat_list[[d]] %notin% names(checktabs[[d]][[y]]))
    return(pat[not_matched])
  })
  names(s) <- station_year[[d]]
  return(s)
})

names(pat_not_matched) <- basename(sdirs)[1:length(checktabs)]


# look into names per station 
dy_names <- lapply(seq(checktabs), function(d){ # for all different stations
  s <- sapply(seq(checktabs[[d]]), function(y){ # for all years available for this station
    names(checktabs[[d]][[y]])
  })
  names(s) <- station_year[[d]]
  return(s)
})
names(dy_names) <- basename(sdirs)[1:length(checktabs)]


```


```{r, warning=FALSE, message=FALSE, eval=F}
# BULL PASS
check_varsel$BullPass # to see whether there are differences over the years
pat_not_matched$BullPass # all good
# dy_names$BullPass$'2012' # all names in respective climate station

# BULL PASS EAST 
check_varsel$BullPassEast # to see whether there are differences over the years
pat_not_matched$BullPassEast # which in pattern could not be matched
# SOLVE: WHY IS MRC_0_cm_deg_C NOT MATCHED ALTHOUGH ITS IN PATTERN AND FILENAMES?????
dy_names$BullPassEast$'2012' # all names in respective climate station

# DON JUAN POND
check_varsel$DonJuanPond # to see whether there are differences over the years
pat_not_matched$DonJuanPond # which in pattern could not be matched
# dy_names$DonJuanPond$'2012' # all names in respective climate station
# adapt pattern to match the station using the data labels word-document 

# GRANITE HARBOUR
check_varsel$GraniteHarbour # to see whether there are differences over the years
pat_not_matched$GraniteHarbour # which in pattern could not be matched
# all good
# dy_names$GraniteHarbour$'2012' # all names in respective climate station

# MARBLE POINT
check_varsel$MarblePoint # to see whether there are differences over the years
pat_not_matched$MarblePoint # which in pattern could not be matched
# dy_names$MarblePoint$'2012' # all names in respective climate station

# MtFlemming
check_varsel$MtFlemming # to see whether there are differences over the years
pat_not_matched$MtFlemming # which in pattern could not be matched
# dy_names$MtFlemming$'2012' # all names in respective climate station

# vICTORIA VALLEY
check_varsel$VictoriaValley # to see whether there are differences over the years
pat_not_matched$VictoriaValley # which in pattern could not be matched
# dy_names$VictoriaValley$'2012' # all names in respective climate station


```


### Now reading in the actual datasets and saving as RMD 

```{r, warning=FALSE, message=FALSE, eval=F}

tabs <- lapply(seq(sdirs), function(d){ # for all different stations
  print(paste0("d= ",d))
  f <- list.files(sdirs[d], full.names=T,pattern=".xls")
  stat_year <- lapply(seq(length(f)), function(y){ # for all years available for this station
  # loop through all of the files
      print(y)
      names_df <- read_excel(f[y],
           col_names = F)
      # get rid of empty columns
      any(is.na(names_df[5,]))
      
      isnadf <- is.na(names_df[200:216,])
      full_cols <- which(colSums(isnadf)!=16)
      
      pastenams <- apply(names_df[1:3,], 2, paste , collapse = "_" )
      pastenams <- gsub(" ", "_", pastenams)
      pastenams <- pastenams[full_cols]
      
      tab <- read_excel(f[y], .name_repair = "unique",
                 skip=3)
      
      isnatabdf <- is.na(tab[200:216,])
      full_cols_tab <- which(colSums(isnatabdf)!=16)
      tab <- tab[,full_cols_tab]
      ncol(tab)
      length(pastenams)
      
      names(tab) <- pastenams
      
      names(tab)[1] <- "ID"
      tab$ID <- shortID[[d]]
      
      return(tab)
  })
})


names(tabs) <- basename(sdirs)
# saveRDS(tabs, here("../calib_valid/AWS/SoilStations/all_SoilStation_data.rds"))


```

Subsetting by the pattern
```{r, warning=FALSE, message=FALSE, eval=F}
# get year corresponding to y per station
station_year <- lapply(seq(tabs), function(d){ # for all different stations
  sapply(seq(tabs[[d]]), function(y){ # for all years available for this station
    tabs[[d]][[y]]$YEAR_NA_NA[1]
  })
})

# subset by name
tabs_varsel <- lapply(seq(tabs), function(d){ # for all different stations
  s <- lapply(seq(tabs[[d]]), function(y){ # for all years available for this station
    sel <- which(names(tabs[[d]][[y]]) %in% pat_list[[d]])
    ts <- tabs[[d]][[y]][,sel]
    return(ts)
  })
  names(s) <- station_year[[d]]
  return(s)
})

names(tabs_varsel) <- basename(sdirs)[1:length(tabs_varsel)]

```

### check, whether all are there
```{r, warning=FALSE, message=FALSE, eval=F}
'%notin%' <- Negate('%in%')
# which columns could not be matched? 
pat_not_matched_tabs <- lapply(seq(tabs_varsel), function(d){ # for all different stations
  s <- lapply(seq(tabs_varsel[[d]]), function(y){ # for all years available for this station
    not_matched <- which(names(tabs_varsel[[d]][[y]])  %notin% pat_list[[d]])
    return(pat_list[[d]][not_matched])
  })
  names(s) <- station_year[[d]]
  return(s)
})

names(pat_not_matched_tabs) <- basename(sdirs)[1:length(tabs_varsel)]

```

All good except for Bull Pass East not having MRC_0_cm_deg_C (don't understand why)
and a lot of missing data in 2019, also Don Juan Pond is missing data in 2019.


Writing to a RDS file 

```{r, warning=FALSE, message=FALSE}
# saveRDS(tabs_varsel, here("../calib_valid/AWS/SoilStations/tabs_varsel.rds"))
# tabs_varsel <- readRDS(here("../calib_valid/AWS/SoilStations/tabs_varsel.rds"))
```

### Match 
```{r, warning=FALSE, message=FALSE, eval=F}
library(dplyr)

cols_per_station <- lapply(seq(tabs_varsel), function(d){
  sapply(seq(tabs_varsel[[d]]), function(y){
  ncol(tabs_varsel[[d]][[y]])
})
})

names(cols_per_station) <- shortID 

BP_one_missing <- do.call("rbind", tabs_varsel$BullPass[c(1,3:length(tabs_varsel$BullPass))])
BPE_no_19 <- do.call("rbind", tabs_varsel$BullPassEast[1:7])
DJP_no_19 <- do.call("rbind", tabs_varsel$DonJuanPond[1:8])

BP_allyears <- bind_rows(BP_one_missing, tabs_varsel$BullPassEast[[2]])
BPE_allyears <- bind_rows(BPE_no_19, tabs_varsel$BullPassEast[[8]])
DJP_allyears <- bind_rows(DJP_no_19, tabs_varsel$DonJuanPond[[9]])
GH_allyears <- do.call("rbind", tabs_varsel$GraniteHarbour)
MP_allyears <- do.call("rbind", tabs_varsel$MarblePoint)
MtF_allyears <- do.call("rbind", tabs_varsel$MtFlemming)
VV_allyears <- do.call("rbind", tabs_varsel$VictoriaValley)


stations <- list(BP_allyears, BPE_allyears, DJP_allyears, GH_allyears, 
     MP_allyears, MtF_allyears, VV_allyears)

names(stations) <- shortID
# saveRDS(stations, here("../calib_valid/AWS/SoilStations/stations_before_unifying.rds"))
```

1. Keep BP, BPE and DJP together for relating soil RH to SMC
2. One full version: kill Soil_Temp_2_cm_White_deg_C and Soil_Temp_2_cm_Black_deg_C as
well as both SOIL_RH_% in BP. Then join all together and rename as in excel file: 
```{r, eval=F, echo=F, message=F, warning=FALSE}
stations$BP[,c(9,10,11,12)] <- NULL
stations$BPE$"Air_Temp_deg_C...6" <- NULL
stations$DJP$"Air_Temp_deg_C...6" <- NULL
stations$MP[,6] <- NULL
stations$BPE$Soil_Temp_2_cm_deg_C <- NULL
stations$BPE$"Soil_RH_2_cm_%" <- NULL
stations$BPE$`Relative_Humidity_%` <- NULL
stations$DJP$`Soil_RH_2_cm_%` <- NULL
stations$BP <- stations$BP[,1:20]

# first Wind Direction_deg_True_N in 7, then Wind Speed in 8
stations$BP <- stations$BP[,c(1:6,8,7,9:20)]
stations$MP <- stations$MP[,c(1:6,8,7,9:20)]
stations$VV <- stations$VV[,c(1:6,8,7,9:20)]

nam <- lapply(seq(stations), function(d){
  names(stations[[d]])
})

stationnames <- data.frame(do.call("cbind", nam))
names(stationnames) <- shortID
stationnames

# new names: 
newnam <- c(
"ID",
"YEAR",
"HOUR",
"DATE",
"AirT",
"SolarRad",
"WindDir",
"WindSpeed",
"SoilT0_2",
"SoilT5_7",
"ER",
"EI",
"Temp",
"ER_COR",
"EI_COR",
"SMC",
"SAL",
"COND",
"COND_COR",
"WAT_COND_COR")

stations_eqal <- lapply(seq(stations), function(d){
  names(stations[[d]]) <- newnam
  return(stations[[d]])
})

# saveRDS(stations_eqal, here("../calib_valid/AWS/SoilStations/stations_allyears.rds"))

all_useful_var <- read_xlsx(here("../calib_valid/AWS/SoilStations/new_vs_old_colnames.xlsx"),sheet=1)
knitr::kable(head(all_useful_var))

var_together <- read_xlsx(here("../calib_valid/AWS/SoilStations/new_vs_old_colnames.xlsx"),sheet=2)
knitr::kable(head(var_together))
```




```{r, eval=F}
stations <- readRDS(here("../calib_valid/AWS/SoilStations/stations_allyears.rds"))
nam <- lapply(seq(stations), function(d){
  names(stations[[d]])
})
allstat <- do.call("rbind", stations)

# saveRDS(allstat, here("../calib_valid/AWS/SoilStations/all_AWSS_all_years.rds"))

```

# The head of the final result
```{r}
saws <- readRDS(here("../calib_valid/AWS/SoilStations/all_AWSS_all_years.rds"))
knitr::kable(head(saws))
```

# observations per climate station
```{r}
table(saws$ID)
```

# log SMC 
```{r}
hist(log(saws$SMC), 
     breaks=60,
     main="log SMC")
```

