---
title: "Calibration and validation data for surface moisture in the McMurdo Dry Valleys"
author: "Maite"
date: "13 9 2021"

output: 
 rmarkdown::html_document:
  #bookdown::html_document2:
    theme: flatly
    highlight: tango
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 6
link-citations: yes
toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning=F, echo = F)
```


```{r, echo=F, message=F, warning=F}
library(sf)
library(mapview)
library(ggplot2)
library(padr)
library(gridExtra)
library(grid)
library(colorspace)
library(here)
library(dplyr)
library(lubridate)
library(readxl)
library(raster)
```

This skript shows calibration and validation data for surface moisture modelling in the McMurdo Dry Valleys in a descriptive way. 

# iButton locations

Format is Degrees Decimal Minutes (DDM), i.e. 77° 30.5' S	164° 45.25' E == Degrees Minutes Seconds (DMS)	77° 30' 29.9988" S	164° 45' 15.0012" E, 
only Taylor Valley comes in DD. 

[this map](https://upload.wikimedia.org/wikipedia/commons/c/c6/McMurdo_Dry_Valleys_from_Antarctic_topographic_reconnaissance_maps_%28USGS-NSF%29.jpg) shows the central Dry Valleys. 

[this map](http://huey.colorado.edu/diatoms/pdf/map_dry_valleys.pdf) shows the whole research area and their names 


Read in iButton data for the different valleys. 

```{r, message=F, results='hide'}

aoi <- st_read(here("../aux_data/Levy_MDV_actually.shp"))

iButDir <- paste0(here("../calib_valid/iButtons/"))

gpkdir <- paste0(iButDir, "/gpkg_DDcorrection/")
f <- list.files(gpkdir,full.names=T)

iB_valley <- lapply(seq(f),function(i){
  st_read(f[[i]])
})


```


Get DEM to write elevation into the location data
```{r, echo=F, message=F, warning=F}
predDir <- here("../predictors/")
pred <- readRDS(list.files(predDir, pattern="full", full.names=T))

pred <- pred[[c("dem", "slope", "aspect",
       "soilraster", "landcoverres")]]
# predcrs <- readRDS(list.files(predDir, pattern="predcrs", full.names=T))

dem <- raster(paste0(predDir, "/DEM_8m_MDV_filled_aoi.tif"))

```

Project iButton data to EPSG 3031
```{r, echo=F, message=F, warning=F}
predcrs <- readRDS(list.files(predDir, pattern="predcrs", full.names=T))
iB_valley <- lapply(seq(iB_valley), function(i){
  x <- st_transform(iB_valley[[i]], crs = predcrs)
  x$Site_Name <- factor(x$Site_Name)
  x
})
names(iB_valley) <- tools::file_path_sans_ext(basename(f))
```


```{r, echo=F, message=F, warning=F}
# How many locations per valley?
loc_p_v <- sapply(seq(iB_valley), function(i){
  length(unique(iB_valley[[i]]$geom))
})
# max = 22 for each unique valley
data.frame("valley"=names(iB_valley), "n_locations"=loc_p_v)

# get locations of unqiue Site_Names
unique_locations <- st_difference(iB_valley$iBut_dec)

pred_loc <- raster::extract(pred, unique_locations)

unique_locations <- cbind(unique_locations, pred_loc)

```



## Display locations
```{r, message=F}

valleys <- unique(unique_locations$Valley)

plots <- lapply(seq(valleys), function(v){
  # cat(valleys[v], ":")
  iBut_Vlly <- unique_locations[unique_locations$Valley==valleys[v],]
  ex <- st_buffer(st_as_sfc(st_bbox(iBut_Vlly)),dist=200)
  DEM_Vlly <- raster::crop(dem,y=st_bbox(ex))
  mapview(DEM_Vlly,na.color="#00000000",
          map.types = "Esri.WorldImagery")+mapview(iBut_Vlly,
                                                   label="Site_Name",zcol="dem")
})
```

Use only data, where temperature is higher than -1°C
```{r, echo=F, message=F, warning=F}

onlySummer <- TRUE 

iButPlots <- lapply(seq(valleys), function(v){
  iBut_Vlly <- iB_valley[[which(names(iB_valley)==valleys[v])]]
  iBut_Vlly <- iBut_Vlly[order(iBut_Vlly$time),]
  # cut into two to not get the gap in the middle where temperatures are too low
  timeslot1 <- iBut_Vlly[order(iBut_Vlly$time) < (max(order(iBut_Vlly$time))/2),] 
  timeslot2 <- iBut_Vlly[order(iBut_Vlly$time) > (max(order(iBut_Vlly$time))/2),] 
  
  iBut_Vlly <- list(timeslot1, timeslot2)
  
  fp <- lapply(seq(iBut_Vlly), function(i){
    
      if(onlySummer == TRUE){
        iBut_Vlly[[i]] <- iBut_Vlly[[i]][month(iBut_Vlly[[i]]$time) %in% c(11,12,1,2),] 
      }
      
      iBut_Vlly[[i]]$Rh[iBut_Vlly[[i]]$T < -1] <- NA
      
      fs <- unique(year(iBut_Vlly[[i]]$time))

    
      pRh <- ggplot(iBut_Vlly[[i]], aes(x=time, y=Rh, color=Site_Name)) +
            scale_colour_viridis_d()+
            geom_line() +
            ggtitle(paste0("field season ", paste(fs, collapse = '/')))+
            xlab("")
    
    
      pT <- ggplot(iBut_Vlly[[i]], aes(x=time, y=T, color=Site_Name)) +
            scale_colour_viridis_d()+
            geom_line() +
            geom_hline(yintercept = -1, linetype = "dotted")+
            xlab("")
      
      fp <- arrangeGrob(pRh, pT, nrow = 2)
  })
  return(fp)  

})

```


To see the iButton logger location names, hold cursor over point.

### All locations at once
There's an interesting cluster with Victoria, McKelvey and Wright Valley that covers a wide range of elevation on a restricted area. Taylor Valley is also interesting due to the amount of available data from there.

Hidden and Alatna Valley are pretty far off and don't really seem to be on open soil areas.  But generally only Taylor and Wright Valley, perhaps Victoria seem to show really open soil where loggers are placed. 
```{r, echo=F, message=F, warning=F}
mapview(dem,na.color="#00000000", map.types = "Esri.WorldImagery")+
  mapview(unique_locations,label="Valley",zcol="dem")
```


### Wright_Valley  
```{r, echo=F, message=F, warning=F}
plots[[2]]
grid.draw(iButPlots[[2]][[1]])
grid.draw(iButPlots[[2]][[2]])

```

### McKelvey_Valley
```{r, echo=F, message=F, warning=F}
plots[[3]]
grid.draw(iButPlots[[3]][[1]])
grid.draw(iButPlots[[3]][[2]])

```

### Victoria_Valley
```{r, echo=F, message=F, warning=F}
plots[[4]]
grid.draw(iButPlots[[4]][[1]])
grid.draw(iButPlots[[4]][[2]])

```

### Alatna_Valley  
```{r, echo=F, message=F, warning=F}
plots[[5]]
grid.draw(iButPlots[[5]][[1]])
grid.draw(iButPlots[[5]][[2]])

```

### Taylor_Valley
```{r, echo=F, message=F, warning=F}
plots[[6]]
grid.draw(iButPlots[[6]][[1]])
grid.draw(iButPlots[[6]][[2]])

```

### Hidden_Valley
```{r, echo=F, message=F, warning=F}
plots[[1]]
grid.draw(iButPlots[[1]][[1]])
grid.draw(iButPlots[[1]][[2]])

```

TV-S-500 seems to be below a snow patch in a shady corner and is fully saturated. 


# Automatic Weather Station Data
### Met Stations can't be used, because Relative Humidity is measured at 3m height. See Met_stations.Rmd for more info. 

### Antarctica Soil Climate Research Stations 
Data can be found [here](https://www.nrcs.usda.gov/wps/portal/nrcs/detail/soils/home/?cid=nrcs142p2_053772). 
There are 9 of them, excluding Minna Bluff and Scott's Base there are: 

1. Bull Pass

established in January of 1999, Wright Valley, 150m elevation, measures  soil temperature to a depth of 120 cm, soil water content (2, 20, 30, 40, 50, and 120 cm depths), air temperature, relative humidity, solar radiation, and wind (speed and direction). Measurements are made at 20-minute intervals and averaged and recorded every hour.

2. Bull Pass East 

established in January of 2012. The station is located in a bowl shaped ridge just east of Bull Pass overlooking the Wright valley in Antarctica. The elevation is about 835 m (2740 ft). This station monitors soil temperature to a depth of 120 cm, soil water content (2, 10, and 25 cm depths), air temperature, relative humidity, soil pore humidity (2 and 10 cm depths), solar radiation, and wind (speed and direction). Measurements are made at 20-minute intervals and averaged and recorded every hour.

3. Don Juan Pond

established in January of 2011. The station is located on a ridge about 700 m (2296 ft) above Don Juan Pond at the upper end of the Wright valley in Antarctica. This station monitors soil temperature to a depth of 120 cm, soil water content (2, 10, and 25 cm depths), air temperature, relative humidity, soil pore humidity (2 and 10 cm depths), solar radiation, and wind (speed and direction). Measurements are made at 20-minute intervals and averaged and recorded every hour.

4. Granite Harbour

established in January of 2003 and is the northern most coastal station. It borders the Ross Sea and is 195 km (120 miles) north of the Minna Bluff station. The elevation is about 4.5 m (15ft). This station monitors soil temperature to a depth of 90 cm, soil water content (2, 5, 15, 25, 45, and 75 cm depths), air temperature, relative humidity, solar radiation, and wind (speed and direction). Measurements are made at 20-minute intervals and averaged and recorded every hour.

5. Marble Point

established in January of 1999 and is located on the coast bordering the Ross Sea. The elevation is 50 m (165 ft). This station monitors soil temperature to a depth of 120 cm, soil water content (2, 20, 50, and 80 cm depths), air temperature, relative humidity, solar radiation, and wind (speed and direction). Measurements are made at 20-minute intervals and averaged and recorded every hour.

6. Mt. Fleming

established in January of 2002 and is the highest at an elevation of 1,700 m (5,600 ft). It was established on Mt. Fleming, at the head of the Wright Valley and adjacent to the polar plateau. The Mt. Fleming station is also located on patterned ground. This station monitors soil temperature to a depth of 60 cm, soil water content (7.5, 15, and 30 cm depths), air temperature, relative humidity, solar radiation, and wind (speed and direction). Measurements are made at 20-minute intervals and averaged and recorded every hour.

7. Victoria Valley

established in January of 1999 and is one of two stations located in the Dry Valleys. The Victoria Valley station is at an elevation of 410 m (1,350 ft) in an area of patterned ground. This station monitors soil temperature to a depth of 120 cm, soil water content (10, 33, 83, 91, and 106 cm depths), air temperature, relative humidity, solar radiation, and wind (speed and direction). Measurements are made at 20-minute intervals and averaged and recorded every hour.


The following map contains information on the location of the Soil Climate Stations with their year of establishment. All of them gather data until 2019. Moreover, the iButton locations are depicted as small points. 


```{r, echo=F, warning=FALSE, message=FALSE}
# ggplot colors come from package scales 
# scales::show_col(hue_pal()(9))

# make some colors to show different calibration and validation sources
library(scales)
cols <- hue_pal()(13)
Scols <- cols[c(2,3,4,5,7,9,11)]
icols <- cols[c(1,6,8,10,12,13)]
```

### Visualization of data acquisition locations and time frames from iButton and AWS

```{r, echo=F, warning=FALSE, message=FALSE}
loc_SoilS <- read.csv2(here("../calib_valid/AWS/SoilStations/Soil_Station_locations.csv"))
loc_SoilS <- loc_SoilS[1:7,]
loc_SoilS$Sread <- as.numeric(loc_SoilS$Sread)
loc_SoilS$Eread <- as.numeric(loc_SoilS$Eread)
loc_SoilS$Nread <- loc_SoilS$Sread*(-1)


SoilS = st_as_sf(loc_SoilS, coords = c("Eread","Nread"), remove = FALSE,
              crs=4326)

SoilS <- SoilS[order(SoilS$station),]

mapview(SoilS, label="station", map.types = "Esri.WorldImagery",
        cex=8, zcol="station", col.regions=Scols)+
   mapview(aoi, map.types = c("Esri.WorldImagery"),
                 alpha.regions = 0,
                 color = 'chartreuse3',
                 lwd=3,legend=FALSE)+
  mapview(unique_locations,label="Valley",zcol="Valley",
          cex=3, 
          col.regions = icols)
  
```



```{r, warning=FALSE, message=FALSE}

timeframe_per_valley <- lapply(seq(valleys), function(v){
    iBut_Vlly <- iB_valley[[which(names(iB_valley)==valleys[v])]]
    return(as.character(c(min(iBut_Vlly$time), max(iBut_Vlly$time))))
})

names(timeframe_per_valley) <- valleys

tf_val <- data.frame(do.call("rbind", timeframe_per_valley))
names(tf_val) <- c("start","end")
tf_val$valley <- rownames(tf_val)
tf_val <- tidyr::gather(tf_val, startend, time, -valley)
tf_val$time <- as.POSIXct(tf_val$time, tz="Pacific/Auckland")

soilfr <- data.frame(paste0(SoilS$station, "_Soil_Stat"), rep("start",7), 
                     as.POSIXct(paste0(SoilS$from, "-01-01 00:01"), format="%Y-%m-%d %H:%M"))
soilto <- data.frame(paste0(SoilS$station, "_Soil_Stat"), rep("end",7), 
                     as.POSIXct(paste0(SoilS$to, "-12-31 00:01"), format="%Y-%m-%d %H:%M"))

names(soilfr) <- c("valley", "startend","time")
names(soilto) <- c("valley", "startend","time")

tf_val <- rbind(tf_val, soilfr, soilto)
tf_val$valley <- as.factor(tf_val$valley)

ggplot(tf_val, aes(x=time, y=valley, color=valley))+
  geom_line(size=3)+
  scale_color_hue(cols)+
  scale_x_datetime(min(tf_val$time), min(tf_val$time), "1 year")+
  ggtitle("SoilStation and iButton deployment timeframes per Valley")+
  theme_minimal()+theme(legend.position = "none",
                        axis.title.y = element_blank(),
                        axis.text.y = element_text(size=12, face="bold"),
                        plot.title = element_text(size = 16, face="bold"),
                        axis.title.x = element_blank(),
                        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

